### 2019-04-03 paper/ecml.tex
Initial Overleaf Import

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Contribution Title\thanks{Supported by organization x.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Princeton University, Princeton NJ 08544, USA \and
Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
\email{lncs@springer.com}\\
\url{http://www.springer.com/gp/computer-science/lncs} \and
ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
\email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The abstract should briefly summarize the contents of the paper in
15--250 words.

\keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
%
%
%
\section{First Section}
\subsection{A Subsection Sample}
Please note that the first paragraph of a section or subsection is
not indented. The first paragraph that follows a table, figure,
equation etc. does not need an indent, either.

Subsequent paragraphs, however, are indented.

\subsubsection{Sample Heading (Third Level)} Only two levels of
headings should be numbered. Lower level headings remain unnumbered;
they are formatted as run-in headings.

\paragraph{Sample Heading (Fourth Level)}
The contribution should contain no more than four levels of
headings. Table~\ref{tab1} gives a summary of all heading levels.

\begin{table}
\caption{Table captions should be placed above the
tables.}\label{tab1}
\begin{tabular}{|l|l|l|}
\hline
Heading level &  Example & Font size and style\\
\hline
Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
\hline
\end{tabular}
\end{table}

\noindent Displayed equations are centered and set on a separate
line.
\begin{equation}
x + y = z
\end{equation}
Please try to avoid rasterized images for line-art diagrams and
schemas. Whenever possible, use vector graphics instead (see
Fig.~\ref{fig1}).

\begin{figure}
%\includegraphics[width=\textwidth]{fig1.eps}
\caption{A figure caption is always placed below the illustration.
Please note that short captions are centered, while long ones are
justified by the macro package automatically.} \label{fig1}
\end{figure}

\begin{theorem}
This is a sample theorem. The run-in heading is set in bold, while
the following text appears in italics. Definitions, lemmas,
propositions, and corollaries are styled the same way.
\end{theorem}
%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%
\begin{proof}
Proofs, examples, and remarks have the initial word in italics,
while the following text appears in normal font.
\end{proof}
For citations of references, we prefer the use of square brackets
and consecutive numbers. Citations using labels or the author/year
convention are also acceptable. The following bibliography provides
a sample reference list with entries for journal
articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
and a homepage~\cite{ref_url1}. Multiple citations are grouped
\cite{ref_article1,ref_lncs1,ref_book1},
\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}
\bibitem{ref_article1}
Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)

\bibitem{ref_lncs1}
Author, F., Author, S.: Title of a proceedings paper. In: Editor,
F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
Springer, Heidelberg (2016). \doi{10.10007/1234567890}

\bibitem{ref_book1}
Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
Location (1999)

\bibitem{ref_proc1}
Author, A.-B.: Contribution title. In: 9th International Proceedings
on Proceedings, pp. 1--2. Publisher, Location (2010)

\bibitem{ref_url1}
LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
Oct 2017
\end{thebibliography}
\end{document}

### 2019-04-03 paper/main.tex
Initial Overleaf Import

\let\origvec\vec
\documentclass[runningheads]{llncs}
\let\springervec\vec
\let\vec\origvec

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\newcommand{\citep}{\cite}
\newcommand{\citet}{\cite}

% Custom packages
\usepackage{graphicx}
%\usepackage{natbib}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{calc}
\usepackage{mathtools}
\usepackage{bbm}
\usepackage{multirow}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{todonotes}
\usepackage{svg}
\usepackage[lofdepth,lotdepth]{subfig}
\usepackage{comment}

\input{mathdef.tex}

\newcommand{\oam}[1]{\todo[inline,color=orange!40]{{\textbf{OM:}~}#1}}

\newcommand{\diff}[1]{\textcolor{ForestGreen}{#1}}

\begin{document}

%
\title{Practical Open-Loop Optimistic Planning}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Edouard Leurent\inst{1,2}\and
Odalric-Ambrym Maillard\inst{1}}
%
\authorrunning{E. Leurent, O-A. Maillard}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{SequeL team, INRIA Lille - Nord Europe, France\\
\email{\{edouard.leurent,odalric.maillard\}@inria.fr}\and
Renault Group, France\\
\email{edouard.leurent@renault.com}}
%
\maketitle              % typeset the header of the contribution
%

\begin{abstract}
    We consider the problem of online planning in a Markov Decision Process when given only access to a generative model, restricted to open-loop policies - i.e. sequences of actions - and under budget constraint. In this setting, the \textit{Open-Loop Optimistic Planning} (\OLOP) algorithm enjoys good theoretical guarantees but is overly conservative in practice, as we show in numerical experiments. We propose a modified version of the algorithm with tighter upper-confidence bounds, \KLOLOP, that leads to better practical performances while retaining the sample complexity bound. Finally, we propose an efficient implementation that significantly improves the time complexity of both algorithms.

    \keywords{Planning \and Online learning \and Tree search.}
\end{abstract}

\section{Introduction}

In a \emph{Markov Decision Process} (MDP), an agent observes its current state $s$ from a state space $S$ and picks an action $a$ from an action space $A$, before transitioning to a next state $s'$ drawn from a transition kernel $\probability{s'|s,a}$ and receiving a bounded reward $r\in[0, 1]$ drawn from a reward kernel $\probability{r|s, a}$. The agent must act so as to optimise its expected cumulative discounted reward $\expectedvalue \sum_t \gamma^t r_t$, also called expected \emph{return}, where $\gamma\in[0,1)$ is the discount factor. In \emph{Online Planning} \cite{Munos2014}, we do not consider that these transition and reward kernels are known as in \emph{Dynamic Programming} \citep{Bellman1957}, but rather only assume access to the MDP through a \emph{generative model} (e.g. a simulator) which yields samples of the next state $s' \sim \probability{s'|s,a}$ and reward $r\sim\probability{r|s, a}$ when queried. Finally, we consider a \emph{fixed-budget} setting where the generative model can only be called a maximum number of times, called the budget $n$.

\emph{Monte-Carlo Tree Search} (\texttt{MCTS}) algorithms were historically motivated by the application of computer Go, and made a first appearance in the CrazyStone software \citet{Coulom2006}. They were later reformulated in the setting of Multi-Armed Bandits by \citet{Kocsis2006} with their \emph{Upper Confidence bounds applied to Trees} (\texttt{UCT}) algorithm. Despite its popularity, \texttt{UCT} has been shown to suffer from several limitations: its sample complexity can be at least doubly-exponential for some problems (e.g. when a narrow optimal path is hidden in a suboptimal branch), which is much worse than uniform planning \citep{Coquelin2007}. The \texttt{Sparse Sampling} algorithm of \citet{Kearns2002} achieves better worst-case performance, but it is still non-polynomial and doesn't adapt to the structure of the MDP. In stark contrast, the \emph{Optimistic Planning for Deterministic systems} (\texttt{OPD}) algorithm considered by \citet{Hren2008} in the case of deterministic transitions and rewards exploits the structure of the cumulative discounted reward to achieve a problem-dependent polynomial bound on sample complexity. A similar line of work in a deterministic setting is that of \texttt{SOOP} and \texttt{OPC} by \cite{Busoniu2013,Busoniu2018} thoush they focus on continuous action spaces. \texttt{OPD} was later extended to stochastic systems with the \emph{Open-Loop Optimistic Planning} (\OLOP) algorithm introduced by \citet{Bubeck2010} in the open-loop setting: we only consider sequences of actions independently of the states that they lead to. This restriction in the space of policies causes a loss of optimality, but greatly simplifies the planning problem in the cases where the state space is large or infinite. More recent work such as \texttt{St0p} \citep{Szorenyi2014} and \texttt{TrailBlazer} \citep{Grill2016} focus on the probably approximately correct (PAC) framework: rather than simply recommending an action to maximise the expected rewards, they return an $\epsilon$-approximation of the value at the root that holds with high probability. This highly demanding framework puts a severe strain on these algorithms that were developed for theoretical analysis only and cannot be applied to real problems.

\paragraph{Contributions} The goal of this paper is to study the practical performances of \OLOP when applied to numerical problems. Indeed, \OLOP was introduced along with a theoretical sample complexity analysis but no experiment was carried-out. Our contribution is threefold:

\begin{itemize}
    \item First, we show that in our experiments \OLOP is overly pessimistic, especially in the low-budget regime, and we provide an intuitive explanation by showcasing an unintended effect that alters the behaviour of \OLOP.
    \item Second, we circumvent this issue by leveraging modern tools from the bandits literature to design and analyse a modified version with tighter upper-confidence bounds called \KLOLOP. We show that we retain the asymptotic regret bounds of $\OLOP$ while improving its performances by an order of magnitude in numerical experiments.
    \item Third, we provide a time and memory efficient implementation of \OLOP and \KLOLOP, bringing an exponential speedup that allows to scale these algorithms to high sample budgets.
\end{itemize}

The paper is structured as follows: in section \ref{sec:kl-olop}, we present \OLOP, give some intuition on its limitations, and introduce \KLOLOP, whose sample complexity is further analysed in section \ref{sec:sample-complexity}. In section \ref{sec:time-complexity}, we propose an efficient implementation of the two algorithms. Finally in section \ref{sec:experiments}, we evaluate them in several numerical experiments.

\subsubsection{Notations}
Throughout the paper, we follow the notations from \citep{Bubeck2010} and use the standard notations over alphabets: a finite word $a \in A^*$ of length $h$ represents a sequence of actions $(a_0, \cdots, a_h) \in A^h$. Its prefix of length $t \leq h$ is denoted $a_{1:t} = (a_0,\cdots,a_t) \in A^t$. $A^\infty$ denotes the set of infinite sequences of actions. Two finite sequences $a\in A^*$ and $b\in A^*$ can be concatenated as $ab\in A^*$, the set of finite and infinite suffixes of $a$ are respectively $a A^* = \{c\in\mathcal{A}^*: \exists b\in A^*$ such that $c=ab\}$ and $aA^\infty$ defined likewise, and the empty sequence is $\emptyset$.

During the planning process, the agent iteratively selects sequences of actions until it reaches the allowed budget of $n$ actions. More precisely, at time $t$ during the $m^{\text{th}}$ sequence, the agent played $a^m_{1:t} = a^m_1 \cdots a^m_t \in A^t$ and receives a reward $Y_t^m$. We denote the probability distribution of this reward as $\nu(a_{1:t}^m) = \probability{Y_t^m | s_t, a^m_t} \prod_{k=1}^{t-1} \probability{s_{k+1} | s_{k}, a_{k}^m}$, and its mean as $\mu(a_{1:t}^m)$, where $s_1$ is the current state.

After this exploration phase, the agent selects an action $a(n)$ so as to minimise the \emph{simple regret} $r_n = V - V(a(n))$, where $V=V(\emptyset)$ and $V(a)$ refers to the value of a sequence of actions $a\in A^h$, that is, the maximum expected discounted cumulative reward one may obtain after executing $a$:

\begin{equation}
\label{eq:value}
V(a) = \sup_{b\in aA^\infty} \sum_{t=1}^\infty \gamma^t\mu(b_{1:t}),
\end{equation}

\section{Kullback-Leibler Open-Loop Optimistic Planning}
\label{sec:kl-olop}

In this section we present \KLOLOP, a combination of the \OLOP algorithm of \citep{Bubeck2010} with the tighter Kullback-Leibler upper confidence bounds from \citep{Cappe2013}. We first frame both algorithms in a common structure before specifying their implementations.

\subsection{General structure}

First, following \OLOP, the total sample budget $n$ is split in $M$ trajectories of length $L$ in the following way: 
\begin{align*}
    &M\text{ is the largest integer such that } M \ceil{\log M/(2 \log 1/\gamma)} \leq n;\\
    &L= \ceil{\log M / (2 \log 1/\gamma)}.
\end{align*}
The look-ahead tree of depth $L$ is denoted $\Tau = \sum_{h=0}^L A^h$.

Then, we introduce some useful definitions. Consider episode $1 \leq m \leq M$. For any $1 \leq h \leq L$ and $a\in A^h$, let

\begin{equation*}
    T_a(m) \eqdef \sum_{s=1}^m \mathbbm{1}\{a^s_{1:h} = a\}
\end{equation*}

\noindent
be the number of times we played an action sequence starting with $a$, and $S_a(m)$ the sum of rewards collected at the last transition of the sequence $a$:

\begin{equation*}
    S_a(m) \eqdef \sum_{s=1}^m Y^s_h \mathbbm{1}\{a^s_{1:h} = a\}
\end{equation*}

\noindent
The empirical mean reward of $a$ is
$\quad\displaystyle{ \hat{\mu}_a(m) \eqdef \frac{S_a(m)}{T_a(m)}} \quad $
if $T_a(m) > 0$, and $+\infty$ otherwise. Here, we provide a more general form for upper and lower confidence bounds on these empirical means:
\begin{eqnarray}
\label{eq:u_mu_a_m}
    U^{\mu}_a(m) &\eqdef \max \left\{q\in I: T_a(m) d(\frac{S_a(m)}{T_a(m)}, q) \leq f(m) \right\}\\
    L^{\mu}_a(m) &\eqdef \min \left\{q\in I: T_a(m) d(\frac{S_a(m)}{T_a(m)}, q) \leq f(m) \right\}
\end{eqnarray}
where $I$ is an interval, $d$ is a divergence on $I\times I \rightarrow \mathbb{R^+}$ and $f$ is a non-decreasing function. They are left unspecified for now and their particular implementations and associated properties will be discussed in the following sections.

These upper-bounds $U^{\mu}_a$ for intermediate rewards finally enable us to define an upper bound $U_a$ for the value $V(a)$ of the entire sequence of actions $a$:

\begin{equation}
\label{eq:Ua}
    U_a(m) \eqdef \sum_{t=1}^h \gamma^t U^{\mu}_{a_{1:t}}(m) + \frac{\gamma^{h+1}}{1-\gamma}
\end{equation}

\noindent
where $\frac{\gamma^{h+1}}{1-\gamma}$ comes from upper-bounding by one every reward-to-go in the sum \eqref{eq:value}, for $t\geq h+1$. In \citep{Bubeck2010}, there is an extra step to "sharpen the bounds" of sequences $a \in A^L$ by taking:

\begin{equation}
\label{eq:Ba}
    B_a(m) \eqdef \inf_{1 \leq t \leq L} U_{a_{1:t}}(m)
\end{equation}

The general algorithm structure is shown in Algorithm \ref{algo:kl-olop}.
We now discuss two specific implementations that differ in their choice of divergence $d$ and non-decreasing function $f$. They are compared in Table \ref{tab:comparison}.

\begin{algorithm}[tp]
\DontPrintSemicolon
\For{each episode $m = 1, \cdots, M$}{
Compute $U_a(m-1)$ from \eqref{eq:Ua} for all $a\in\Tau$\;
Compute $B_a(m-1)$ from \eqref{eq:Ba} for all $a\in A^L$\;\label{alg:b_values_compute}
Sample a sequence with highest B-value: $a^m \in \argmax_{a\in A^L} B_a(m-1)$\;
}
\Return the most played sequence $a(n) \in \argmax_{a\in A^L} T_a(M)$
\caption{General structure for Open-Loop Optimistic Planning}
\label{algo:kl-olop}
\end{algorithm}

\begin{table}[tp]
    \caption{Different implementations of Algorithm \ref{algo:kl-olop} in \OLOP and \KLOLOP}
    \label{tab:comparison}
    \centering
    \begin{tabular}{ccc}
    \toprule
        Algorithm & \OLOP & \KLOLOP \\
        \midrule
        Interval $I$ & $\mathbb{R}$ & [0, 1] \\
        Divergence $d$ & $\dquad$ & $\dber$ \\
        $f(m)$ & $4 \log M$ & $2\log M + 2 \log\log M$\\
        \bottomrule
    \end{tabular}
\end{table}

\subsection{OLOP}
\label{sec:kl-olop-olop}
To recover the original \OLOP algorithm of \citet{Bubeck2010} from Algorithm \ref{algo:kl-olop}, we can use a quadratic divergence $\dquad$ on $I=\mathbb{R}$ and a constant function $f_4$ defined as follows:
\begin{equation*}
    \dquad(p,q) \eqdef 2(p-q)^2,\qquad
    f_4(m) \eqdef 4 \log M
\end{equation*}
Indeed, in this case $U^{\mu}_a(m)$ can then be explicitly computed as:
\begin{align*}
    U^{\mu}_a(m) &= \max \left\{q\in \mathbb{R}: 2(\frac{S_a(m)}{T_a(m)} - q)^2 \leq \frac{4 \log M }{T_a(m)} \right\} = \hat{\mu}_a(m) + \sqrt{\frac{2 \log M}{T_a(m)}}
\end{align*}
which is the Chernoff-Hoeffding bound used originally in section 3.1 of \cite{Bubeck2010}.

\subsection{An unintended behaviour}

From the definition of $U_a(m)$ as an upper-bound of the value of the sequence $a$, we expect increasing sequences $(a_{1:t})_t$ to have non-increasing upper-bounds. Indeed, every new action $a_t$ encountered along the sequence is a potential loss of optimality.
However, this property is only true if the upper-bound defined in \eqref{eq:u_mu_a_m} belongs to the reward interval $[0,1]$.

\begin{lemma}(Monotony of $U_a(m)$ along a sequence)
\label{lemma:seq_values}

\begin{itemize}
    \item If it holds that $U^{\mu}_b(m) \in [0, 1]$ for all $b\in A^*$, then for any $a\in A^L$ the sequence $(U_{a_{1:h}}(m))_{1\leq h \leq L}$ is non-increasing, and we simply have $B_a(m) = U_a(m)$.
    \item Conversely, if $U^{\mu}_b(m) > 1$ for all $b\in A^*$, then for any $a\in A^L$ the sequence $(U_{a_{1:h}}(m))_{1\leq h \leq L}$ is non-decreasing, and we have $B_a(m) = U_{a_{1:1}}(m)$.
\end{itemize}
\end{lemma}

\begin{proof}
We prove the first proposition, and the same reasoning applies to the second. For $a\in A^L$ and $1 \leq h \leq L - 1$, we have by \eqref{eq:Ua}:

\begin{align*}
    U_{a_{1:h+1}}(m) - U_{a_{1:h}}(m) &= \gamma^{h+1}U^{\mu}_{a_{1:h+1}}(m) + \frac{\gamma^{h+2}}{1-\gamma} - \frac{\gamma^{h+1}}{1-\gamma}\\
    &= \gamma^{h+1}(\underbrace{U^{\mu}_{a_{1:h+1}}(m)}_{\in [0, 1]} - 1) \leq 0
    %&\leq 0
\end{align*}

\noindent
We can conclude that $(U_{a_{1:h}}(m))_{1\leq h \leq L}$ is non-increasing and that $B_a(m) = \inf_{1 \leq h \leq L} U_{a_{1:h}}(m) = U_{a_{1:L}}(m) = U_a(m)$.
\qed
\end{proof}

Yet, the Chernoff-Hoeffding bounds used in \OLOP start in the $U^{\mu}_a(m) > 1$ regime -- initially $U^{\mu}_a(m) = \infty$ -- and can remain in this regime for a long time especially in the near-optimal branches where $\hat{\mu}_a(m)$ is close to one.

Under these circumstances, the Lemma \ref{lemma:seq_values} has a drastic effect on the search behaviour. Indeed, as long as a subtree under the root verifies $U^{\mu}_a(m) > 1$ for every sequence $a$, then all these sequences share the same B-value $B_a(m) = U_{a_{1:1}(m)}$. This means that \OLOP cannot differentiate them and exploit information from their shared history as intended, and behaves as uniform sampling instead.
Once the early depths have been explored sufficiently, \OLOP resumes its intended behaviour, but the problem is only shifted to deeper unexplored subtrees.

 This consideration motivates us to leverage the recent developments in the Multi-Armed Bandits literature, and modify the upper-confidence bounds for the expected rewards $U^\mu_a(m)$ so that they respect the reward bounds.

\subsection{KL-OLOP}
\label{sec:kl-olop-kl-olop}

\noindent
We propose a novel implementation of Algorithm \ref{algo:kl-olop} where we leverage the analysis of the kl-UCB algorithm from \citep{Cappe2013} for multi-armed bandits with general bounded rewards.
Likewise, we use the Bernoulli Kullback-Leibler divergence defined on the interval $I=[0,1]$ by:

\begin{equation*}
    \dber(p, q) \eqdef p \log \frac{p}{q} + (1-p)\log\frac{1-p}{1-q}
\end{equation*}

\noindent
with, by convention, $0 \log 0 = 0 \log 0/0 = 0$ and $x \log x /0 = +\infty$ for $x>0$.
This divergence and the corresponding bounds are illustrated in Figure \ref{fig:ukl}.

\begin{figure}[tp]
    \centering
    \includegraphics[width=0.6\textwidth]{img/ukl.eps}
    \caption{The Bernoulli Kullback-Leibler divergence $\dber$, and the corresponding upper and lower confidence bounds $U^{\mu}_a$ and $L^{\mu}_a$ for the empirical average $\hat{\mu_a}$. Lower values of $f(m)$ give tighter confidence bounds that hold with lower probabilities.}
    \label{fig:ukl}
\end{figure}

$U^{\mu}_a(m)$ and $L^{\mu}_a(m)$ can be efficiently computed using Newton iterations, as for any $p\in[0, 1]$ the function $q \rightarrow \dber(p,q)$ is strictly convex and increasing (resp. decreasing) on the interval [p, 1] (resp. [0, p]).

Moreover, we use the constant function $f_2: m \rightarrow 2 \log M + 2 \log\log M$. This choice is justified in the end of section \ref{sec:regret-proof}. Because $f_2$ is lower than $f_4$, the Figure \ref{fig:ukl} shows that the bounds are tighter and hence less conservative than that of \OLOP, which should increase the performance, provided that their associated probability of violation does not invalidate the regret bound of \OLOP.

\begin{remark}[Upper bounds sharpening]
\label{rmk:sharpen}
The introduction of the B-values $B_a(m)$ was made necessary in \OLOP by the use of Chernoff-Hoeffding confidence bounds which are not guaranteed to belong to [0, 1]. On the contrary, we have in \KLOLOP that $U^\mu_a(m) \in I = [0,1]$ by construction. By Lemma \ref{lemma:seq_values}, the upper bounds sharpening step in line \ref{alg:b_values_compute} of Algorithm \ref{algo:kl-olop} is now superfluous as we trivially have $B_a(m) = U_a(m)$ for all $a\in A^L$.
\end{remark}

\section{Sample complexity}
\label{sec:sample-complexity}

We say that $u_n = \tilde{O}(v_n)$ if there exist $\alpha, \beta >0$ such that $u_n \leq \alpha \log(v_n)^\beta v_n$.
Let us denote the proportion of near-optimal nodes $\kappa_2$ as:

\begin{equation*}
    \label{eq:kappa}
    \kappa_2 = \limsup_{h\rightarrow\infty}{\left|\left\{a\in a^H:V(a) \geq V - 2\frac{\gamma^{h+1}}{1-\gamma}\right\}\right|^{1/h}}
\end{equation*}

\begin{theorem}[Sample complexity]
\label{thm:regret}
We show that \KLOLOP enjoys the same asymptotic regret bounds as \OLOP. More precisely, for any $\kappa' > \kappa_2$, \KLOLOP satisfies:

\begin{equation*}
    \expectedvalue r_n = \begin{cases}
      \tilde{0}\left(n^{-\frac{\log 1/\gamma}{\log \kappa'}}\right), & \text{if}\ \gamma\sqrt{\kappa'} > 1 \\
      \tilde{0}\left(n^{-\frac{1}{2}}\right), & \text{if}\ \gamma\sqrt{\kappa'} \leq 1
    \end{cases}
\end{equation*}
\end{theorem}

\section{Time and memory complexity}
\label{sec:time-complexity}

After having considered the sample efficiency of \OLOP and \KLOLOP, we now turn to study their time and memory complexities. We will only mention the case of \KLOLOP for ease of presentation, but all results easily extend to \OLOP.

The Algorithm \ref{algo:kl-olop} requires, at each episode, to compute and store in memory of the reward upper-bounds and U-values of all nodes in the tree $\Tau = \sum_{h=0}^L A^h$.
Hence, its time and memory complexities are 
\begin{equation}
    C(\KLOLOP) = O(M|\Tau|) = O(MK^L).
\end{equation}

The curse of dimensionality brought by the branching factor $K$ and horizon $L$ makes it intractable in practice to actually run \KLOLOP in its original form even for small problems. However, most of this computation and memory usage is wasted, as with reasonable sample budgets $n$ the vast majority of the tree $\Tau$ will not be actually explored and hence does not hold any valuable information.

We propose in Algorithm \ref{algo:lazy-kl-olop} a lazy version of \KLOLOP which only stores and processes the explored subtree, as shown in Figure \ref{fig:tree}, while preserving the inner workings of the original algorithm.

\begin{figure}[ht]
    \centering
    \includesvg[width=0.6\textwidth]{img/tree}
    \caption{A representation of the tree $\Tau_m^+$, with $K = 2$ actions and after episode $m = 2$, when two sequences have been sampled. They are represented with solid lines and dots \textbullet, and they constitute the explored subtree $\Tau_m$. When extending $\Tau_m$ with the missing children of each node, represented with dashed lines and diamonds $\diamond$, we obtain the full extended subtree $\Tau_m^+$. The set of its leaves is denoted $\LL_m^+$ and shown as a dotted set.}
    \label{fig:tree}
\end{figure}

\begin{algorithm}[tp]
\DontPrintSemicolon
Let $M$ be the largest integer such that $M \log M/(2 \log 1/\gamma) \leq n$\;
Let $L = \log M / (2 \log 1/\gamma)$\;
Let $\Tau_0^+ = \LL_0^+ = \{\emptyset\}$\;
\For{each episode $m = 1, \cdots, M$}{
Compute $U_a(m-1)$ from \eqref{eq:Ua} for all $a\in\Tau_{m-1}^+$\;
Compute $B_a(m-1)$ from \eqref{eq:Ba} for all $a\in \LL_{m-1}^+$\;
Sample a sequence with highest B-value: $a \in \argmax_{a\in \LL_{m-1}^+} B_a(m-1)$\;
Choose an arbitrary continuation $a^m \in aA^{L-|a|}$\tcp*{e.g. uniformly}
Let $\Tau_m^+ = \Tau_{m-1}^+$ and $\LL_m^+ = \LL_{m-1}^+$\;
\For{$t=1, \cdots, L$}{
    \If{$a^m_{1:t} \not \in \Tau_{m}^+$}{
    Add $a^m_{1:t-1}A$ to $\Tau_{m}^+$ and $\LL_{m}^+$\;
    Remove $a^m_{1:t-1}$ from $\LL_{m}^+$
    }
}
}
\Return the most played sequence $a(n) \in \argmax_{a\in \LL_m^+} T_a(M)$
\caption{Lazy Open Loop Optimistic Planning}
\label{algo:lazy-kl-olop}
\end{algorithm}

\begin{theorem}[Consistency]
\label{thm:consistency}
The set of sequences returned by Algorithm \ref{algo:lazy-kl-olop} is the same as the one returned by Algorithm \ref{algo:kl-olop}.
In particular, Algorithm \ref{algo:lazy-kl-olop} enjoys the same regret bounds as in Theorem \ref{thm:regret}.
\end{theorem}

\begin{property}[Time and memory complexity]
Algorithm \ref{algo:lazy-kl-olop} has time and memory complexities of:
\begin{equation*}
    C(\texttt{Lazy KL-OLOP}) = O(KLM^2)
\end{equation*}

The corresponding complexity gain compared to the original Algorithm \ref{algo:kl-olop} is: 
\begin{equation*}
    \frac{C(\texttt{Lazy KL-OLOP})}{C(\KLOLOP)} = \frac{n}{K^{L-1}}
\end{equation*}
which highlights that only a subtree corresponding to the sample budget $n$ is processed instead of the search whole tree $\Tau$.
\end{property}
\begin{proof}
At episode $m = 1, \cdots, M$, we compute and store in memory of the reward upper-bounds and U-values of all nodes in the subtree $\Tau_m^+$. Moreover, the tree $\Tau_m^+$ is constructed iteratively by adding K nodes at most L times at each episode from 0 to $m$. Hence, $|\Tau_m^+| = O(mKL)$.
This yields directly $C(\texttt{Lazy KL-OLOP}) = \sum_{m=1}^M O(mKL) = O(M^2KL)$.
\qed
\end{proof}

\section{Proof of Theorem \ref{thm:regret}}
\label{sec:regret-proof}

We follow step-by step the pyramidal proof of \citep{Bubeck2010}, and adapt it to the Kullback-Leibler upper confidence bound. The adjustments resulting from the change of confidence bounds are \diff{highlighted}. The proofs of lemmas which are not significantly altered are listed in the Supplementary Material.

We start by recalling their notations.
Let $1 \leq H \leq L$ and $a^* \in A^L$ such that $V(a^*) = V$.
Considering sequences of actions of length $1 \leq h \leq H$, we define the subset $\mathcal{I}_h$ of near-optimal sequences and the subset $\mathcal{J}$ of sub-optimal sequences that were near-optimal at depth $h-1$:
\begin{equation*}
    \mathcal{I}_h = \left\{a \in A^h: V - V(a) \leq 2\frac{\gamma^{h+1}}{1-\gamma}\right\}, \mathcal{J}_h = \left\{a \in A^h: a_{1:h-1} \in \mathcal{I}_{h-1} \text{ and } a \not\in \mathcal{I}_h\right\}
\end{equation*}

By convention, $\mathcal{I}_0 = \{\emptyset\}$. From the definition of $\kappa_2$, we have that for any $\kappa'>\kappa_2$, there exists a constant C such that for any $h \geq 1$,
\begin{equation*}
    |\mathcal{I}_h| \leq C {\kappa'}^h
\end{equation*}
Hence, we also have $|\mathcal{J}_h| \leq K|\mathcal{I}_{h-1}| = O({\kappa'}^h)$.

Now, for $1\leq m \leq M$, $a \in A^t$ with $t \leq h$, $h'<h$, we define the set $\mathcal{P}^a_{h,h'}(m)$ of suffixes of $a$ in $\mathcal{J}_h$ that have been played at least a certain number of times:
\begin{equation*}
    \mathcal{P}^a_{h,h'}(m) = \left\{ b\in a A^{h-t}\cap \mathcal{J}_h : T_b(m) \geq \diff{2f(m)}(h+1)^2\gamma^{2(h'-h+1)} + 1 \right\}
\end{equation*}

and the random variable:

\begin{equation*}
    \tau^a_{h,h'}(m) = \mathbbm{1}\{T_a(m-1) < \diff{2f(m)}(h+1)^2\gamma^{2(h'-h+1)} + 1 \leq T_a(m)\}
\end{equation*}

\begin{lemma}[Regret and sub-optimal pulls]
\label{lemma:expected-regret}
The following holds true:
\begin{equation*}
    r_n \leq \frac{2K \gamma^{H+1}}{1-\gamma} +\frac{3K}{M}\sum_{h=1}^H\sum_{a\in\mathcal{J}_h}\frac{\gamma^h}{1-\gamma}T_a(M)
\end{equation*}
\end{lemma}

The rest of the proof is devoted to the analysis of the term $\expectedvalue \sum_{a\in \mathcal{J}_h} T_a(M)$. The next lemma describes under which circumstances a suboptimal sequence of actions in $\mathcal{J}_h$ can be selected.

\begin{lemma}[Conditions for sub-optimal pull]
\label{lemma:sub-optimal-pull}
Assume that at step $m+1$ we select a sub-optimal sequence $a^{m+1}$: there exist $0 \leq h \leq L,  a\in \mathcal{J}_h$ such that $a^{m+1} \in aA^*$. Then, it implies that one of the following propositions is true:
\begin{equation}
\label{eq:cond-ukl}
\tag{UCB violation}
    \diff{U_{a^*}(m)} < V,
\end{equation}
or
\begin{equation}
\tag{LCB violation}
\label{eq:cond-lkl}
    \sum_{t=1}^h \gamma^t \diff{L^{\mu}_{a_{1:t}}(m)} \geq V(a),
\end{equation}
or
\begin{equation}
\tag{Large CI}
\label{eq:cond-dkl}
    \sum_{t=1}^h \gamma^t\diff{\left(U^{\mu}_{a_{1:t}}(m) - L^{\mu}_{a_{1:t}}(m)\right)} > \frac{\gamma^{h+1}}{1-\gamma}
\end{equation}
\end{lemma}
\begin{proof}
As $a^{m+1}_{1:h} = a$ and \diff{because the U-values are monotonically increasing along sequences of actions} (see Remark \ref{rmk:sharpen} and Lemma \ref{lemma:seq_values}), we have $U_a(m) \geq U_{a^{m+1}}(m)$. Moreover, by Algorithm \ref{algo:kl-olop}, we have $a^{m+1} = \argmax_{a \in A^L}  U_a(m)$ and $a^*\in A^L$, so $U_{a^{m+1}}(m) \geq U_{a^*}(m)$ and finally $U_a(m) \geq U_{a^*}(m)$.

Assume that \eqref{eq:cond-ukl} is false, then:
\begin{equation}
\label{eq:ukl-verifie}
    \sum_{t=1}^h \gamma^t U^{\mu}_{a_{1:t}}(m) + \frac{\gamma^{h+1}}{1-\gamma} = U_a(m) \geq U_{a^*}(m) \geq V
\end{equation}
Assume that \eqref{eq:cond-lkl} is false, then:
\begin{equation}
\label{eq:lkl-verifie}
    \sum_{t=1}^h \gamma^t L^{\mu}_{a_{1:t}}(m) < V(a),
\end{equation}
By taking the difference \eqref{eq:ukl-verifie} - \eqref{eq:lkl-verifie}, 
\begin{equation*}
    \sum_{t=1}^h \gamma^t \left(U^{\mu}_{a_{1:t}}(m) - L^{\mu}_{a_{1:t}}(m)\right) + \frac{\gamma^{h+1}}{1-\gamma} > V - V(a)
\end{equation*}
But $a \in \mathcal{J}_h$, so $V - V(a) \geq \frac{2\gamma^{h+1}}{1-\gamma}$, which yields \eqref{eq:cond-dkl} and concludes the proof.
\qed
\end{proof}

\diff{In the following lemma, for each episode $m$ we bound the probability of \eqref{eq:cond-ukl} or \eqref{eq:cond-lkl} by a desired confidence level $\delta_m$, whose choice we postpone until the end of this proof. For now, we simply assume that we picked a function $f$ that satisfies $f(m)\log (m) e^{-f(m)} = O(\delta_m)$. We also denote $\Delta_M = \sum_{m=1}^{M}\delta_m$.}

\begin{lemma}[Boundary crossing probability]
\label{lemma:boundary-crossing-prob}
The following holds true, for any $1 \leq h \leq L$ and $m \leq M$,
\begin{equation*}
    \probability{\text{\eqref{eq:cond-ukl} or \eqref{eq:cond-lkl} is true}} = \diff{O((L+h)\delta_m)}
\end{equation*}
\end{lemma}
\begin{proof}
Since $V \leq \sum_{t=1}^h \gamma^t \mu(a^*_{1:t}) + \frac{\gamma^{h+1}}{1-\gamma}$, we have,
\begin{align*}
   \probability{\eqref{eq:cond-ukl}} &=  \probability{U_{a^*}(m) \leq V}\\
    &= \probability{\sum_{t=1}^L \gamma^t U^{\mu}_{a^*_{1:t}}(m) \leq \sum_{t=1}^L \gamma^t \mu(a^*_{1:t})}\\
    &\leq \probability{\exists 1\leq t \leq L : U^{\mu}_{a^*_{1:t}}(m) \leq \mu(a^*_{1:t})} \\
    &\leq \sum_{t=1}^L\probability{U^{\mu}_{a^*_{1:t}}(m) \leq \mu(a^*_{1:t})}
\end{align*}

In order to bound this quantity, \diff{we reduce the question to the application of a deviation inequality}. For all $1\leq t\leq L$, we have on the event $\{U^{\mu}_{a^*_{1:t}}(m) \leq \mu(a^*_{1:t})\}$ that $\hat{\mu}_{a^*_{1:t}}(m) \leq U^{\mu}_{a^*_{1:t}}(m) \leq \mu(a^*_{1:t}) < 1$. Therefore, for all $0 < \delta < 1 - \mu(a^*_{1:t})$, by definition of $U^{\mu}_{a^*_{1:t}}(m)$:

\begin{equation*}
    d(\hat{\mu}_{a^*_{1:t}}(m), U^{\mu}_{a^*_{1:t}}(m)+\delta) > \frac{f(m)}{T_{a^*_{1:t}}(m)}
\end{equation*}

As $d$ is continuous on $(0,1)\times[0, 1]$, we have by letting $\delta \leftarrow 0$ that:

\begin{equation*}
    d(\hat{\mu}_{a^*_{1:t}}(m), U^{\mu}_{a^*_{1:t}}(m)) \geq \frac{f(m)}{T_{a^*_{1:t}}(m)}
\end{equation*}

Since d is non-decreasing on $[\hat{\mu}_{a^*_{1:t}}(m), \mu(a^*_{1:t})]$,

\begin{equation*}
    d(\hat{\mu}_{a^*_{1:t}}(m), \mu(a^*_{1:t})) \geq d(\hat{\mu}_{a^*_{1:t}}(m), U^{\mu}_{a^*_{1:t}}(m)) \geq \frac{f(m)}{T_{a^*_{1:t}}(m)}
\end{equation*}

We have thus shown the following inclusion:

\begin{equation*}
    \{U^{\mu}_{a^*_{1:t}}(m) \leq \mu(a^*_{1:t})\} \subseteq \left\{ \mu(a^*_{1:t}) > \hat{\mu}_{a^*_{1:t}}(m) \text{ and } d(\hat{\mu}_{a^*_{1:t}}(m), \mu(a^*_{1:t})) \geq \frac{f(m)}{T_{a^*_{1:t}}(m)} \right\}
\end{equation*}

Decomposing according to the values of $T_{a^*_{1:t}}(m)$ yields:

\begin{equation*}
    \{U^{\mu}_{a^*_{1:t}}(m) \leq \mu(a^*_{1:t})\} \subseteq \bigcup_{n=1}^m \left\{ \mu(a^*_{1:t}) > \hat{\mu}_{a^*_{1:t}, n} \text{ and } d(\hat{\mu}_{a^*_{1:t}, n}, \mu(a^*_{1:t})) \geq \frac{f(m)}{n} \right\}
\end{equation*}

\diff{We now apply the deviation inequality} provided in Lemma 2 of Appendix A in \citep{Cappe2013}: $\forall \epsilon > 1$, provided that $0 < \mu(a^*_{1:t}) < 1$,

\begin{equation*}
\probability{\bigcup_{n=1}^m \left\{ \mu(a^*_{1:t}) > \hat{\mu}_{a^*_{1:t}, n} \text{ and } n \dber(\hat{\mu}_{a^*_{1:t}, n}, \mu(a^*_{1:t})) \geq \epsilon \right\}} \leq e\ceil{\epsilon \log m}e^{-\epsilon}\,.
\end{equation*}

By choosing $\epsilon = f(m)$, it comes
\begin{align*}
    \probability{\eqref{eq:cond-ukl}} &\leq \sum_{t=1}^L \diff{e\ceil{f(m)\log m}e^{-f(m)}} = \diff{O(L\delta_m)}
\end{align*}

The same reasoning gives: $\quad\displaystyle{
    \probability{\eqref{eq:cond-lkl}} = \diff{O(h\delta_m)}}$.
\qed
\end{proof}

\begin{lemma}[Confidence interval length and number of plays]
\label{lemma:ci-length}
Let $1 \leq h \leq L$, $a\in \mathcal{J}_h$ and $0 \leq h' < h$. Then  \eqref{eq:cond-dkl} is not satisfied if the following propositions are true:
\begin{equation}
\label{eq:sampled-enough-h}
   \forall 0\leq t\leq h', T_{a_{1:t}}(m) \geq \diff{2f(m)}(h+1)^2\gamma^{2(t-h-1)}
\end{equation}
and
\begin{equation}
\label{eq:sampled-enough}
   T_{a}(m) \geq \diff{2f(m)}(h+1)^2\gamma^{2(h'-h-1)}
\end{equation}
\end{lemma}
\begin{proof}
We start by providing an explicit upper-bound for the length of the confidence interval $U^{\mu}_{a_{1:t}} - L^{\mu}_{a_{1:t}}$. By Pinsker's inequality:

\begin{equation*}
    \diff{\dber(p, q) > \dquad(p, q)}
\end{equation*}

Hence for all $C>0$, 
\begin{equation*}
    \dber(p, q) \leq C   \implies 2(q - p)^2 < C  \implies p - \sqrt{C/2} < q < p + \sqrt{C/2}
\end{equation*}
And thus, for all $b\in A^*$, by definition of $U^{\mu}$ and $L^{\mu}$:
\begin{align*}
    U^{\mu}_{b}(m) - L^{\mu}_{b}(m) &\leq \frac{S_b(m)}{T_b(m)} + \sqrt{\frac{f(m)}{2T_b(m)}} -  \left(\frac{S_b(m)}{T_b(m)} - \sqrt{\frac{f(m)}{2T_b(m)}}\right) 
    = \diff{\sqrt{\frac{2f(m)}{T_b(m)}}}
\end{align*}

Now, assume that \eqref{eq:sampled-enough-h} and \eqref{eq:sampled-enough} are true. Then, we clearly have:
\begin{align*}
    \sum_{t=1}^h \gamma^t\left(U^{\mu}_{a_{1:t}}(m) - L^{\mu}_{a_{1:t}}(m)\right) &\leq \sum_{t=1}^{h'} \gamma^t \sqrt{\frac{2f(m)}{T_{a_{1:t}}(m)}} + \sum_{t=h'+1}^h \gamma^t \sqrt{\frac{2f(m)}{T_{a_{1:t}}(m)}} \\
    &\leq \frac{1}{(h+1)\gamma^{-h-1}} \sum_{t=1}^{h'} 1 + \frac{1}{(h+1)\gamma^{-h-1}} \sum_{t=h'+1}^h \gamma^{t-h'}  \\
    &\leq \frac{\gamma^{h+1}}{h+1} \left( h' + \frac{\gamma}{1-\gamma} \right)\leq \frac{\gamma^{h+1}}{1-\gamma}\,.\qquad\qquad\qed
\end{align*}
\end{proof}

\begin{lemma}
\label{lemma:size_Ph}
Let $1 \leq h \leq L,  a\in \mathcal{J}_h$ and $0\leq h'<h$. Then $\tau^a_{h,h'}=1$ implies that either equation \eqref{eq:cond-ukl} or \eqref{eq:cond-lkl} is satisfied or the following proposition is true:

\begin{equation}
    \label{eq:P-min-size}
    \exists 1\leq t \leq h': |\mathcal{P}_{h,h'}^{a_{1:t}}(m)| < \gamma^{2(t-h')}
\end{equation}
\end{lemma}

\begin{lemma}
\label{lemma:expected-P-size}
Let $1\leq h\leq L$ and $0 \leq h' < h$. Then the following holds true,

\begin{equation*}
    \expectedvalue |\mathcal{P}^\emptyset_{h,h'}(M)| = \tilde{O}\left(\gamma^{-2h'}\mathbbm{1}_{h'>0}\sum_{t=0}^{h'}(\gamma^2 \kappa')^t + (\kappa')^h\diff{\Delta_M} \right).
\end{equation*}
\end{lemma}

\begin{lemma}
\label{lemma:expected-plays-count}
Let $1\leq h\leq L$. The following holds true,
\begin{equation*}
    \expectedvalue \sum_{a\in \mathcal{J}_h} T_a(M) = \tilde{O}\left(\gamma^{-2h} + \diff{(\kappa')^h(1+M\Delta_M+\Delta_M) + (\kappa'\gamma^{-2})^h\Delta_M}\right)
\end{equation*}
\end{lemma}

Thus by combining Lemma \ref{lemma:expected-regret} and \ref{lemma:expected-plays-count} we obtain:
\begin{equation*}
    \expectedvalue r_n = \tilde{O}\left(\gamma^H + \gamma^{-H}M^{-1} + (\kappa' \gamma)^{H}M^{-1}\diff{(1+M\Delta_M+\Delta_M)} + (\kappa')^H\gamma^{-H}M^{-1}\diff{\Delta_M}\right)
\end{equation*}
Finally,
\begin{itemize}
    \item if $\kappa'\gamma^2 \leq 1$, we take $H = \floor{\log M / (2\log 1/\gamma)}$ to obtain:
    \begin{equation*}
        \expectedvalue r_n = \tilde{O}\left(M^{-\frac{1}{2}} + M^{-\frac{1}{2}} + M^{-\frac{1}{2}}M^{\frac{\log \kappa'}{2\log 1/\gamma}} \diff{\Delta_M}\right)
    \end{equation*}
    For the last term to be of the same order of the others, we need to have $\Delta_M = O(M^{-\frac{\log \kappa'}{2\log 1/\gamma}})$. Since $\kappa'\gamma^2 \leq 1$, this can be achieved by taking \diff{$\Delta_M = O(M^{-1})$}.
    \item if $\kappa'\gamma^2 > 1$, we take $H = \floor{\log M / \log \kappa'}$ to obtain:
    \begin{equation*}
        \expectedvalue r_n = \tilde{O}\left(M^{\frac{\log \gamma}{\log \kappa'}} + M^{\frac{\log \gamma}{\log \kappa'}}\diff{(1+M\Delta_M+\Delta_M)} + M^{\frac{\log 1/\gamma}{\log \kappa'}}\diff{\Delta_M}\right)
    \end{equation*}
    Since $\kappa'\gamma^2 > 1$, the dominant term in this sum is $M^{\frac{\log \gamma}{\log \kappa'}}M\Delta_M$. Again, taking \diff{$\Delta_M = O(M^{-1})$} yields the claimed bounds.
    The dominant terms are 
\end{itemize}
Thus, the claimed bounds are obtained in both cases as long as we can impose $\Delta_M = O(M^{-1})$, that is, find a sequence $(\delta_m)_{1\leq m\leq M}$ and a function $f$ verifying:
\begin{equation}
    \sum_{m=1}^M \delta_m = O(M^{-1})\quad \text{and}\quad f(m)\log (m) e^{-f(m)} = O(\delta_m)
\end{equation}

By choosing \diff{$\delta_m = M^{-2}$ and $f(m) = 2 \log M + 2 \log\log M$}, the corresponding \KLOLOP algorithm does achieve the regret bound claimed in Theorem \ref{thm:regret}.

\section{Experiments}
\label{sec:experiments}

We have performed some numerical experiments to evaluate and compare the following planning algorithms\footnote[1]{The source code is available at \url{https://eleurent.github.io/kl-olop/}}:
\begin{itemize}
    \item \texttt{Random}: returns an action at random. We use it as a baseline for minimal performance.
    \item \texttt{OPD}: the \emph{Optimistic Planning for Deterministic systems} from \citep{Hren2008}, used as a baseline of optimal performance. This planner is only suited for deterministic environments, and exploits this property to obtain faster rates. However, it is expected to fail in stochastic environments.
    \item \OLOP: as described in section \ref{sec:kl-olop-olop}.\footnote[2]{Note that we use the lazy version of $\OLOP$ and $\KLOLOP$ presented in Section \ref{sec:time-complexity}, otherwise the exponential running-time would have been prohibitive.}
    \item \KLOLOP: as described in section \ref{sec:kl-olop-kl-olop}.\footnotemark[2]
    \item \texttt{KL-OLOP}(1): an aggressive version of \KLOLOP where we used $f_1(m) = \log M$ instead of $f_2(m)$. This threshold function makes the upper bounds even tighter, at the cost of an increase probability of violation. Hence, we expect this solution to be more efficient in close-to-deterministic environments. However, since we have no theoretical guarantee concerning its regret as we do with $\KLOLOP$, it might not be conservative enough and converge too early to a suboptimal sequence, especially in highly stochastic environments.
\end{itemize}

They are evaluated by solving the following tasks, using a discount factor of $\gamma=0.8$:
\begin{itemize}
    \item A \href{https://github.com/eleurent/highway-env/}{highway driving} environment \citep{Leurent2018}: a vehicle is driving on a road randomly populated with other slower drivers, and must make their way as fast as possible while avoiding collisions by choosing on the the following actions: \texttt{faster}, \texttt{slower}, \texttt{change-lane-left}, \texttt{change-lane-right}, \texttt{no-op}.
    \item A \href{https://github.com/maximecb/gym-minigrid}{gridworld} environment \citep{gym_minigrid}: the agent navigates in a randomly-generated gridworld composed of either empty cells, terminal lava cells, and goal cells where a reward of $1$ is collected at the first visit.
    \item A stochastic version of the gridworld environment with noisy rewards, where the noise is modelled as a Bernoulli distribution with a 15\% probability of error, i.e. receiving a reward of 1 in an empty cell or 0 in a goal cell.
\end{itemize}

\begin{figure}[htp]
\centering
\begin{minipage}{\textwidth}
    \subfloat[Highway]{
      \includesvg[width=\textwidth]{img/hw_return.svg}
      \label{sub:highway}
    }
    \newline
    \subfloat[Gridworld]{
      \includesvg[width=0.5\textwidth]{img/gw_return.svg}
      \label{sub:gridworld}
    }
    \subfloat[Stochastic gridworld]{
      \includesvg[width=0.5\textwidth]{img/gw_stoch_return.svg}
      \label{sub:gridworld_stoch}
    }

    \caption{Numerical experiments: for each environment-agent configuration, we compute the average return over 100 runs –- along with its 95\% confidence interval –- with respect to the available budget $n$.}
    \label{fig:experiments}
\end{minipage}
\end{figure}

The results of our experiments are shown in Figure \ref{fig:experiments}. The \texttt{ODP} algorithm converges very quickly to the optimal return in the two first environments shown in Figure \ref{sub:highway} and Figure \ref{sub:gridworld}, because it exploits their deterministic nature: it needs neither to estimate the rewards through upper-confidence bounds nor to sample whole sequences all the way from the root when expanding a leaf, which provides a significant speedup. It can be seen of an oracle allowing to measure the conservativeness of stochastic planning algorithms. And indeed, even before introducing stochasticity, we can see that \OLOP performs quite badly on the two environments, only managing to solve them with a budget in the order of $10^{3.5}$. In stark contrast, \KLOLOP makes a much better use of its samples and reaches the same performance an order of magnitude faster. Moreover, in the stochastic gridworld environment shown in Figure \ref{sub:gridworld_stoch}, we observe that the deterministic \texttt{ODP} planner's performance saturates as it settles to suboptimal trajectories, as expected. Conversely, the stochastic planners all find better-performing open-loop policies, which justifies the need for this framework. Again, \KLOLOP converges an order of magnitude faster than \OLOP. Finally, \texttt{KL-OLOP}(1) enjoys good performance overall and displays the most satisfying trade-off between aggressiveness in deterministic environments and conservativeness in stochastic environments; hence we recommend this tuning for practical use.

\section{Conclusion}

We introduced an enhanced version of the \OLOP algorithm for open-loop online planning, whose design was motivated by an investigation of the over-conservative search behaviours of \OLOP. We analysed its sample complexity and showed that the original regret bounds are preserved, while its empirical performances are increased by an order of magnitude in several numerical experiments. Finally, we proposed an efficient implementation that benefits from a substantial speedup, facilitating its use for real-time planning applications.

\bibliographystyle{splncs04}
\bibliography{references}
%\begin{comment}
\appendix

\section{Supplementary proofs of Theorem \ref{thm:regret}}

\subsection{Lemma \ref{lemma:expected-regret}}
\begin{proof}
See proof of Lemma 4 in \citep{Bubeck2010}.
\qed
\end{proof}

\subsection{Lemma \ref{lemma:size_Ph}}
\begin{proof}
The event $\tau^a_{h,h'}=1$ implies $a^{m+1}\in a A^*$ and \eqref{eq:sampled-enough}. This implies by Lemma \ref{lemma:sub-optimal-pull} that either \eqref{eq:cond-ukl}, \eqref{eq:cond-lkl} or \eqref{eq:cond-dkl} is satisfied. Now by Lemma \ref{lemma:ci-length} this implies that either \eqref{eq:cond-ukl} is true or \eqref{eq:cond-lkl} is true or \eqref{eq:sampled-enough-h} is false. We now prove that if \eqref{eq:P-min-size} is not satisfied then \eqref{eq:sampled-enough} is true, which clearly ends the proof.
This follows from: For any $0 \leq t \leq h'$:

\begin{align*}
    T_{a_{1:t}}(m) &= \sum_{b\in a_{1:t}A^{h-t}} T_b(m) \geq \sum_{b\in \mathcal{P}^{a_{1:t}}_{h,h'}} T_b(m) \\
    &\geq \left(\gamma^{2(t-h')}\right) \left(2f(m)(h+1)^2\gamma^{2(h'-h-1)}\right)\\
    &= 2f(m)(h+1)^2\gamma^{2(t-h-1)}\,.\qquad\qquad\qed
\end{align*}
%\
\end{proof}

\subsection{Lemma \ref{lemma:expected-P-size}}

\begin{proof}
The proof is identical to that of Lemma 9 in \citep{Bubeck2010}, where they show:

\begin{equation*}
    |\mathcal{P}^\emptyset_{h,h'}(M)| \leq \mathbbm{1}_{h'>0}\sum_{t=0}^{h'}(\gamma^{2(t-h')}|\mathcal{I}_t| + \sum_{m=1}^{M}\sum_{a\in \mathcal{J}_h} \mathbbm{1}\{\eqref{eq:cond-ukl}\text{ or }\eqref{eq:cond-lkl}\}
\end{equation*}
Taking the expectation and applying Lemma \ref{lemma:boundary-crossing-prob} yield the claimed bound.
\qed
\end{proof}

\subsection{Lemma \ref{lemma:expected-plays-count}}

\begin{proof}
The proof is identical to that of Lemma 10 in \citep{Bubeck2010}, where they show:
\begin{equation*}
    \sum_{a\in \mathcal{J}_h} T_a(M) = \tilde{O}\left((\kappa')^h + \gamma^{-2h}\sum_{h'=1}^{h-1} \gamma^{2h'}|\mathcal{P}^\emptyset_{h,h'}(M)| + M|\mathcal{P}^\emptyset_{h,0}(M)|\right)
\end{equation*}
Taking the expectation and applying the bound of Lemma \ref{lemma:expected-P-size} give the claimed bound.
\qed
\end{proof}

\section{Proof of Theorem \ref{thm:consistency}}

\begin{proof}

To prove consistency of Algorithm \ref{algo:lazy-kl-olop}, we need to show that the sequences of actions $a^m$ sampled at every episode are chosen arbitrarily from the same sets as in Algorithm \ref{algo:lazy-kl-olop}.
Namely,

\begin{equation*}
    \left\{ b\in a A^{L-|a|} : a \in \argmax_{a\in \LL_{m-1}^+} B_a(m-1)\right\} = \argmax_{a\in A^L} B_a(m-1)
\end{equation*}

To that end, we first introduce some useful notations:

\begin{paragraph}{Definition}
% We remind that $\Tau$ refers to the whole search tree $\Tau = \sum_{h=0}^L A^h$.
Let $\Tau_m$ be the set of visited nodes after episode $m$:
\begin{equation*}
    \Tau_m \eqdef \left\{a\in A^*: T_a(m) > 0\right\}
\end{equation*}

We also define its extension $\Tau_m^+$ of visited nodes and their children:
\begin{equation*}
    \Tau_m^+ \eqdef \Tau_m + \Tau_m A
\end{equation*}

Now for $a\in A^*$, $\pi_m(a)$ (resp. $\pi_m^+(a)$) refers to its longest prefix within $\Tau_m$ (resp. $\Tau_m^+$):
\begin{align*}
    \pi_m(a) &\eqdef \argmax_{b\in\Tau_m} \{|b|: a\in b A^* \} \\
    \pi_m^+(a) &\eqdef \argmax_{b\in\Tau_m^+} \{|b|: a\in b A^* \}
\end{align*}

Finally, $\LL_m$ and $\LL_m^+$ are the image of $A^L$ by $\pi_m$ and $\pi_m^+$, respectively.
\begin{align*}
    \LL_m &\eqdef \pi_m(A^L) \\
    \LL_m^+ &\eqdef \pi_m^+(A^L) \}
\end{align*}
\end{paragraph}

\begin{remark}[About children extensions]
We could frame Algorithm \ref{algo:lazy-kl-olop} in terms of $\Tau_m$ and $\LL_m$, for which mathematical proofs are more straight-forward. However, the iterative construction of $\LL_m$ is tricky and it would require inverting $\pi_m$ on $\LL_m$ which is non-trivial. On the contrary, introducing their extensions  $\Tau_m^+$ and $\LL_m^+$ slightly complicates the proof, but greatly simplifies the construction of $\LL_m^+$ and the computation of ${\pi_m^+}^{-1}$ on $\LL_m^+$, which is why we use these sets in practice.
\end{remark}

\begin{lemma}[Sets construction]
$\Tau_m^+$ and $\LL_m^+$ are indeed the sets computed in Algorithm  \ref{algo:lazy-kl-olop}.
\end{lemma}
\begin{proof}
Note that for each episode $1 \leq m \leq M - 1$, we have:
\begin{equation}
\label{eq:tau_mp1}
    \Tau_{m+1} = \Tau_{m} + \sum_{t=0}^L a^{m+1}_{1:t}
\end{equation}
Indeed, the nodes visited at least once at time $m+1$ where either already visited once at time $m$ (e.g. in $\Tau_{m}$) or have been visited for the first time during episode $m+1$, which means they are a prefix of $a^{m+1}$. The reverse is clearly true as well.

This enables to write:
\begin{align*}
    \Tau_{m+1}^+ &= \Tau_{m+1} + \Tau_{m+1}A & \text{by definition}\\
    &= \Tau_{m} + \sum_{t=0}^L a^{m+1}_{1:t} + (\Tau_{m} + \sum_{t=0}^L a^{m+1}_{1:t})A & \text{by \eqref{eq:tau_mp1}} & \\
    &= (\Tau_{m} + \Tau_{m}A) + \sum_{t=0}^L a^{m+1}_{1:t} + \sum_{t=0}^L a^{m+1}_{1:t}A & \\
    &= \Tau_{m}^+ + a^{m+1}_{1:0} + \sum_{t=0}^L a^{m+1}_{1:t}A &\text{ as } \sum_{t=1}^L a^{m+1}_{1:t} \subset \sum_{t=0}^L a^{m+1}_{1:t}A\\
    &=  \Tau_{m}^+ + \sum_{t=0}^L a^{m+1}_{1:t}A  &\text{ as }a^{m+1}_{1:0}=\emptyset \in \Tau_{0} \subset \Tau_{m} \subset \Tau_m^+
\end{align*}
This recursion is the one implemented in Algorithm \ref{algo:lazy-kl-olop}: at each episode $m$, we add to $\Tau_{m}^+$ the children of the nodes along the sampled action sequence $a^{m}$.

Finally, we highlight that $\LL_m^+ = \pi^+(A^L)$ is the set of leaves of $\Tau_{m}^+$.
Indeed, nodes of $\LL_m^+$ belong to $\Tau_{m}^+$, but they cannot have a child in $\Tau_{m}^+$ as it would contradict the definition of $\LL_m^+$. Conversely, any leaf $a$ of $\Tau_{m}^+$ can be continued arbitrarily to a sequence $b$ of $A^L$, which  $a = \pi_m^+(b) \in \pi^+(A^L) = \LL_m^+$.

Thus, when updating $\Tau_{m-1}^+$, the set of its leaves is updated accordingly: when the children of a leaf $a^m_{1:t-1}$ are added to $\Tau_m^+$, they become new leaves in place of their parent. Hence, they are added to $\LL_m^+$ while $a^m_{1:t-1}$ is removed from it.
\qed
\end{proof}

\begin{lemma}[U-values conservation]
\label{lemma:value-conservation}
For all $a \in A^*$,
\begin{equation*}
    U_a(m) = U_{\pi_m(a)}(m) = U_{\pi_m^+(a)}(m)
\end{equation*}
\end{lemma}
\begin{proof}
Let $a \in A^*$, denote $h=|a|$ and $h'=|\pi_m(a)|$.

By definition of $\pi_m(a)$, $0 \leq h' \leq h$, and
\begin{itemize}
    \item for $1\leq t \leq h'$, we have $a_{1:t} = {\pi_m(a)}_{1:t}$ ;
    \item for $h'+1\leq t \leq h$, we have $a_{1:t} \not \in \Tau_m$, hence $T_{a_{1:t}}(m) = 0$ and $U^{\mu}_{a_{1:t}}(m) = 1$.
\end{itemize}
Then,
\begin{align*}
    U_a(m) &= \sum_{t=1}^h \gamma^t U^{\mu}_{a_{1:t}}(m) + \frac{\gamma^{h+1}}{1-\gamma} \\
    &= \sum_{t=1}^{h'} \gamma^t U^{\mu}_{a_{1:t}}(m) + \sum_{t=h'+1}^h \gamma^t \underbrace{U^{\mu}_{a_{1:t}}(m)}_1 + \frac{\gamma^{h+1}}{1-\gamma} \\
    &= \sum_{t=1}^{h'} \gamma^t U^{\mu}_{{\pi_m(a)}_{1:t}}(m) + \frac{\gamma^{h'+1}}{1-\gamma} \\
    &= U_{\pi_m(a)}(m)
\end{align*}

Now, consider $\pi_m^+(a) \in \Tau_m^+$.
By definition, it belongs either to $\Tau_m$ or $\Tau_m A$.
\begin{itemize}
    \item If $\pi_m^+(a) \in \Tau_m$, then $\pi_m^+(a) = \pi_m(a)$ and $U_{\pi_m^+(a)}(m) = U_{\pi_m(a)}(m)$.
    \item Else, $\pi_m^+(a) \in \Tau_m A$ and $p(\pi_m^+(a)) = \pi_m(a)$.

    As $\pi_m^+(a) \not \in \Tau_m$, we have $T_{\pi_m^+(a)}(m) = 0$ and $U^{\mu}_{\pi_m^+(a)}(m) = 1$.
    This yields:

    \begin{equation*}
        U_{\pi_m^+(a)}(m) = \sum_{t=1}^{h'} \gamma^t U^{\mu}_{{\pi_m^+(a)}_{1:t}}(m) + \gamma^{h'+1} \underbrace{U^{\mu}_{{\pi_m^+(a)}}(m)}_1 + \frac{\gamma^{h'+2}}{1-\gamma} = U_{\pi_m(a)}(m)
    \end{equation*}

\end{itemize}
We showed that $U_{\pi_m^+(a)}(m) = U_{\pi_m(a)}(m)$, which concludes the proof.
\qed
\end{proof}

\begin{lemma}[Inverse projection]
\label{lemma:inverse-proj}
For all $a\in \LL_m^+$ of length $h\leq L$,
\begin{equation*}
    {\pi_m^+}^{-1}(a) = a A^{L-h}
\end{equation*}

This allows to easily pick a sequence inside ${\pi_m^+}^{-1}(a)$: just continue the sequence $a$ with a default action of $A$ (e.g. the first) until it reaches length $L$.
\end{lemma}
\begin{proof}
Let $a\in \LL_m^+$.

By definition of $\pi_m^+$, any sequence in ${\pi_m^+}^{-1}(a)$ is a suffix of $a$ of length $L$, so we clearly have the direct inclusion ${\pi_m^+}^{-1}(a) \subset a A^{L-h}$.

Now for the other side: let $b\in a A^{L-h}$, i.e. $a=b_{1:h}$. We need to show that $\pi_m^+(b) = a$.
As $a\in \LL_m^+$, there exists $c\in A^L$ such that $\pi_m^+(c) = a$.
\begin{itemize}
    \item If h = L, then $b=a$, so $b \in \LL_m^+ \subset \Tau_m^+$, and hence $\pi_m^+(b)=b=a$.
    \item If h < L, we can show by contradiction that $a \not \in \Tau_m$. Indeed, if $a \in \Tau_m$, then $c_{1:h+1}$ is the child of a node of $\Tau_m$ and hence belongs to $\Tau_m^+$. But then, $c_{1:h+1}$ is a prefix of $c$ in $\Tau_m^+$ with greater length than $a$, which contradicts the definition of $a = \pi_m^+(c)$.

    Now, because $a \not \in \Tau_m$, it is also true for all suffixes of $a$, and in particular for $b_{1:t}$ with $h \leq t \leq L$. Indeed, we have $a^s_{1:t} = b_{1:t} \implies a^s_{1:h} = b_{1:h} = a$, so:
    \begin{equation*}
        T_{b_{1:t}}(m) = \sum_{s=1}^m \mathbbm{1}\{a^s_{1:t} = b_{1:t}\} \leq \sum_{s=1}^m \mathbbm{1}\{a^s_{1:h} = a\} = T_a(m) = 0
    \end{equation*}
    Hence, $b_{1:t} \not \in \Tau_m$ for all $h \leq t \leq L$, so in particular $b_{1:t} \not \in \Tau_m^+$ for all $h+1 \leq t \leq L$. Since $b_{1:h} = a \in \Tau_m^+$, $a$ is indeed the longest prefix of $b$ in $\Tau_m^+$, that is: $\pi_m^+(b) = a$.
\end{itemize}
We have shown the other side of the inclusion: $a A^{L-h} \subset {\pi_m^+}^{-1}(a)$, which entails that the two sets are in fact equal.
\qed
\end{proof}

We can now conclude our proof of Theorem \ref{thm:consistency}: at episode $m$, \KLOLOP samples a sequence of action $a^m$ within the set $\argmax_{a\in A^L} U_a(m)$. 
However, we have:

\begin{align*}
    \argmax_{c\in A^L} U_c(m) &= \argmax_{c\in A^L} U_{\pi_m^+(c)}(m) & \text{by Lemma \ref{lemma:value-conservation}} \\
    &= {\pi_m^+}^{-1}\left(\argmax_{a\in {\pi_m^+}(A^L)} U_{a}(m)\right) & \\
    &= \left\{b\in{\pi_m^+}^{-1}(a) : a\in \argmax_{a\in \LL_m^+} U_{a}(m)\right\} & \\
    &= \left\{b\in a A^{L-|a|} : a\in \argmax_{a\in \LL_m^+} U_{a}(m)\right\} & \text{by Lemma \ref{lemma:inverse-proj}} 
\end{align*}

Thus, at each episode the sequence of actions $a^m$ sampled by Algorithm \ref{algo:lazy-kl-olop} could have been sampled by Algorithm \ref{algo:kl-olop} as well.

In particular, if the arbitrary rule used to pick a sequence from a set is the same for the two algorithms, then the sampled sequences $a^m$ will be identical, will have the same visit count $T_{a^m}(m)$, and in the end the returned action $a(n)$ will be the same.
\qed
\end{proof}
%\end{comment}
\end{document}

### 2019-04-03 paper/plot_kl.py

@@ -1,0 +1,52 @@
+import numpy as np
+import matplotlib.pyplot as plt
+# import seaborn as sns
+# sns.set()
+from rl_agents.utils import bernoulli_kullback_leibler
+
+vkl = lambda p, q: p * np.log(p/q) + (1 - p)*np.log((1-p)/(1-q))
+kl = bernoulli_kullback_leibler
+diff_ukl = lambda S, N, f, v: np.array([N*kl(S/N, q) - f for q in v])
+
+def ukl(S, N, f, count=100):
+    xx = np.linspace(0, 1, count)
+    dd = diff_ukl(S, N, f, xx)
+    i = np.nonzero(dd < 0)[0][-1]
+    return xx[i]
+
+def lkl(S, N, f, count=100):
+    xx = np.linspace(0, 1, count)
+    dd = diff_ukl(S, N, f, xx)
+    i = np.nonzero(dd < 0)[0][1]
+    return xx[i]
+
+
+mu = 0.4
+N = 10
+S = mu*N
+f = 0.4*N  # compare to KL = f/N
+
+
+plt.rc('text', usetex=True)
+plt.rc('font', family='serif', size=12)
+
+
+fig = plt.figure()
+ax = fig.add_subplot(1, 1, 1)
+x = np.linspace(0, 1, 1000)
+plt.text(mu, f/N, r'$\frac{1}{T_a}f(m)$', rotation=0, horizontalalignment="center", verticalalignment="bottom")
+plt.plot(x, vkl(S/N, x), label=r"$d(\hat{\mu}_a, q)$")
+plt.xlim([0, 1])
+plt.ylim([0, 1.5])
+plt.legend()
+xticks = [0, lkl(S, N, f), mu, ukl(S, N, f), 1]
+plt.xticks(xticks, ('0', r'$L^{\mu}_a$', r'$\hat{\mu}_a$', r'$U^{\mu}_a$', '1'))
+plt.yticks([])
+xticks_grid = [lkl(S, N, f), ukl(S, N, f)]
+yticks_grid = [0, f / N]
+ax.set_xticks(xticks_grid, minor=True)
+ax.set_yticks(yticks_grid, minor=True)
+ax.grid(which='minor')
+plt.savefig("img/ukl.eps")
+plt.show()
+

### 2019-04-05 paper/main.tex
Updates from Overleaf

@@
 \usepackage{calc}
+\usepackage{bbm}
 \usepackage{mathtools}
@@
 \begin{itemize}
-    \item First, we show that in our experiments \OLOP is overly pessimistic, especially in the low-budget regime, and we provide an intuitive explanation by showcasing an unintended effect that alters the behaviour of \OLOP.
+    \item First, we show that in our experiments \OLOP is overly pessimistic, especially in the low-budget regime, and we provide an intuitive explanation by casting light on an unintended effect that alters the behaviour of \OLOP.
     \item Second, we circumvent this issue by leveraging modern tools from the bandits literature to design and analyse a modified version with tighter upper-confidence bounds called \KLOLOP.
@@
 \subsection{An unintended behaviour}
-
+\label{sec:kl-olop-behaviour}
 From the definition of $U_a(m)$ as an upper-bound of the value of the sequence $a$, we expect increasing sequences $(a_{1:t})_t$ to have non-increasing upper-bounds.
@@
     For the last term to be of the same order of the others, we need to have $\Delta_M = O(M^{-\frac{\log \kappa'}{2\log 1/\gamma}})$.
-Since $\kappa'\gamma^2 \leq 1$, this can be achieved by taking \diff{$\Delta_M = O(M^{-1})$}.
+Since $\kappa'\gamma^2 \leq 1$, we achieve this by taking \diff{$\Delta_M = O(M^{-1})$}.
     \item if $\kappa'\gamma^2 > 1$, we take $H = \floor{\log M / \log \kappa'}$ to obtain:
@@
     Since $\kappa'\gamma^2 > 1$, the dominant term in this sum is $M^{\frac{\log \gamma}{\log \kappa'}}M\Delta_M$. Again, taking \diff{$\Delta_M = O(M^{-1})$} yields the claimed bounds.
-    The dominant terms are 
 \end{itemize}
@@
 \begin{itemize}
-    \item \texttt{Random}: returns an action at random. We use it as a baseline for minimal performance.
+    \item \texttt{Random}: returns an action at random, we use it as a minimal performance baseline.
     \item \texttt{OPD}: the \emph{Optimistic Planning for Deterministic systems} from \citep{Hren2008}, used as a baseline of optimal performance.
@@
 
-They are evaluated by solving the following tasks, using a discount factor of $\gamma=0.8$:
+They are evaluated on the following tasks, using a discount factor of $\gamma=0.8$:
 \begin{itemize}
-    \item A \href{https://github.com/eleurent/highway-env/}{highway driving} environment \citep{Leurent2018}: a vehicle is driving on a road randomly populated with other slower drivers, and must make their way as fast as possible while avoiding collisions by choosing on the the following actions: \texttt{faster}, \texttt{slower}, \texttt{change-lane-left}, \texttt{change-lane-right}, \texttt{no-op}.
+    \item A \href{https://github.com/eleurent/highway-env/}{highway driving} environment \citep{Leurent2018}: a vehicle is driving on a road randomly populated with other slower drivers, and must make their way as fast as possible while avoiding collisions by choosing on the the following actions: \texttt{change-lane-left}, \texttt{change-lane-right}, \texttt{no-op}, \texttt{faster}, \texttt{slower}.
     \item A \href{https://github.com/maximecb/gym-minigrid}{gridworld} environment \citep{gym_minigrid}: the agent navigates in a randomly-generated gridworld composed of either empty cells, terminal lava cells, and goal cells where a reward of $1$ is collected at the first visit.
@@
 In stark contrast, \KLOLOP makes a much better use of its samples and reaches the same performance an order of magnitude faster.
-Moreover, in the stochastic gridworld environment shown in Figure \ref{sub:gridworld_stoch}, we observe that the deterministic \texttt{ODP} planner's performance saturates as it settles to suboptimal trajectories, as expected.
+Examples of expanded trees are showcased in the Supplementary Material.
+Furthermore, in the stochastic gridworld environment shown in Figure \ref{sub:gridworld_stoch}, we observe that the deterministic \texttt{ODP} planner's performance saturates as it settles to suboptimal trajectories, as expected.
 Conversely, the stochastic planners all find better-performing open-loop policies, which justifies the need for this framework.
@@
 %\begin{comment}
+\newpage
 \appendix
+\begin{center}
+    \Large Supplementary Material
+\end{center}
+
+\section{Examples of expanded trees}
+
+The trees expanded by different planning algorithms in the Highway environment are displayed in Figure \ref{fig:trees}.
+They were all generated in the same initial root state, under the same budget of $n=10^3$ calls to the generative model.
+We observe that the tree developed by the \OLOP algorithm is quite evenly balanced, which suggests that \OLOP behaves as uniform planning as hypothesised in \ref{sec:kl-olop-behaviour}.
+On the contrary, \KLOLOP is much more efficient and produced a highly unbalanced tree, with most of the sampled sequences densely concentrated around the optimal trajectory, while the other branches under the root have barely been sampled more than once.
+\ODP is even more efficient in this deterministic setting, as it does not need to run the trajectories from the root down to depth $L$, nor to explore twice the same transition.
+
+\begin{figure}[pht]
+    \centering
+    \includesvg[width=0.64\textwidth]{img/OLOP}
+    \includesvg[width=0.64\textwidth]{img/KL-OLOP}
+    \includesvg[width=0.64\textwidth]{img/ODP}
+    \caption{The look-ahead trees expanded by different planning algorithms for a budget $n=10^3$, shown down to depth 6. The width of edges is proportional to the nodes visit count $T_a(M)$.}
+    \label{fig:trees}
+\end{figure}
 
@@
 \begin{proof}
-See proof of Lemma 4 in \citep{Bubeck2010}.
+The proof is identical to that of Lemma 4 in \citep{Bubeck2010}.
+Since $\arg \max _{a \in A} T_{a}(M)$, and $\sum_{a \in A} T_{a}(M)=M,$ we have $T_{a(n)}(M) \geq M / K$, and thus:
+\begin{equation*}
+\frac{M}{K}(V-V(a(n))) \leq(V-V(a(n))) T_{a(n)}(M) \leq \sum_{m=1}^{M} V-V\left(a^{m}\right)
+\end{equation*}
+Hence, we have, $r_{n} \leq \frac{K}{M} \sum_{m=1}^{M} V-V\left(a^{m}\right)$. Now remark that, for any sequence of actions $a\in A^L$, we have either:
+\begin{itemize}
+    \item $a_{1 : H} \in \mathcal{I}_{H} ;$ which implies $V-V(a) \leq \frac{2 \gamma^{H+1}}{1-\gamma}$
+    \item or there exists $1\leq h \leq H$ such that $a_{1:h} \in \mathcal{J}_h$; which implies $V-V(a) \leq V-V\left(a_{1 : h-1}\right)+\frac{\gamma^{h}}{1-\gamma} \leq \frac{3 \gamma^{h}}{1-\gamma}$.
+\end{itemize}
+Thus we can write:
+\begin{equation*}
+\begin{aligned} \sum_{m=1}^{M}\left(V-V\left(a^{m}\right)\right) &=\sum_{m=1}^{M}\left(V-V\left(a^{m}\right)\right)\left(\mathbbm{1}\left\{a^{m} \in \mathcal{I}_{H}\right\}+\mathbbm{1}\left\{\exists 1 \leq h \leq H : a_{1 : h}^{m} \in \mathcal{J}_{h}\right\}\right) \\ & \leq \frac{2 \gamma^{H+1}}{1-\gamma} M+3 \sum_{h=1}^{H} \sum_{a \in \mathcal{J}_{h}} \frac{\gamma^{h}}{1-\gamma} T_{a}(M) \end{aligned}
+\end{equation*}
 \qed
@@
 \begin{proof}
-The proof is identical to that of Lemma 9 in \citep{Bubeck2010}, where they show:
+The proof is identical to that of Lemma 9 in \citep{Bubeck2010}.
+
+\noindent
+Let $h'\geq 1$ and $0 \leq s \leq h'$. We introduce the following random variables:
+\begin{equation*}
+m_{s}^{a}=\min \left(M, \min \left\{m \geq 0 :\left|\mathcal{P}_{h, h^{\prime}}^{a}(m)\right| \geq \gamma^{2\left(s-h^{\prime}\right)}\right\}\right).
+\end{equation*}
+We will prove recursively that,
+\begin{equation}
+\label{eq:toprove}
+\left|\mathcal{P}_{h, h^{\prime}}^{\emptyset}(m)\right| \leq \sum_{t=0}^{s} \gamma^{2\left(t-h^{\prime}\right)}\left|\mathcal{I}_{t}\right|+\sum_{a \in \mathcal{I}_{s}}\left|\mathcal{P}_{h, h^{\prime}}^{a} \setminus \cup_{t=0}^{s} \mathcal{P}_{h, h^{\prime}}^{a_{1:t}}\left(m_{t}^{a_{1:t}}\right)\right|
+\end{equation}
+The result is true for $s = 0$ since $\mathcal{I}_0 = \{\emptyset\}$ and by definition of $m^\emptyset_0$,
+\begin{equation*}
+\left|\mathcal{P}_{h, h^{\prime}}^{\emptyset}(m)\right| \leq \gamma^{-2 h^{\prime}}+\left|\mathcal{P}_{h, h^{\prime}}^{\emptyset}(m) \setminus \mathcal{P}_{h, h^{\prime}}^{\emptyset}\left(m_{0}^{\emptyset}\right)\right|
+\end{equation*}
+Now let us assume that the result is true for $s<h'$. We have:
+\begin{align*}
+\sum_{a \in \mathcal{I}_{s}}\left|\mathcal{P}_{h, h^{\prime}}^{a}(m) \setminus \cup_{h, h^{\prime}}^{a_{1 : t}}\left(m_{t}^{a_{1 : t}}\right)\right|&=\sum_{a \in \mathcal{I}_{s+1}}\left|\mathcal{P}_{h, h^{\prime}}^{a}(m) \setminus \cup_{t=0}^{s} \mathcal{P}_{h, h^{\prime}}^{a_{1 : t}}\left(m_{t}^{a_{1 : t}}\right)\right|\\
+&\leq \sum_{a \in \mathcal{I}_{s+1}} \gamma^{2\left(s+1-h^{\prime}\right)}+\left|\mathcal{P}_{h, h^{\prime}}^{a}(m) \setminus \cup_{t=0}^{s+1} \mathcal{P}_{h, h^{\prime}}^{a_{1 : t}}\left(m_{t}^{a_{1 : t}}\right)\right|\\
+&= \gamma^{2\left(s+1-h^{\prime}\right)}\left|\mathcal{I}_{s+1}\right|+\sum_{a \in \mathcal{I}_{s+1}}\left|\mathcal{P}_{h, h^{\prime}}^{a}(m) \setminus \cup_{t=0}^{s+1} \mathcal{P}_{h, h^{\prime}}^{a_{1 ; t}}\left(m_{t}^{a_{1 : t}}\right)\right|
+\end{align*}
+which ends the proof of \eqref{eq:toprove}. Thus we proved (by taking $s=h'$ and $m=M$):
+\begin{equation*}
+\begin{aligned}\left|\mathcal{P}_{h, h^{\prime}}^{\emptyset}(M)\right| & \leq \sum_{t=0}^{h^{\prime}} \gamma^{2\left(t-h^{\prime}\right)}\left|\mathcal{I}_{t}\right|+\sum_{a \in \mathcal{I}_{h^{\prime}}}\left|\mathcal{P}_{h, h^{\prime}}^{a}(M) \setminus \cup_{t=0}^{s+1} \mathcal{P}_{h, h^{\prime}}^{a_{1 : t}}\left(m_{t}^{a_{1 : t}}\right)\right.\\ &=\sum_{t=0}^{h^{\prime}} \gamma^{2\left(t-h^{\prime}\right)}\left|\mathcal{I}_{t}\right|+\sum_{a \in \mathcal{J}_{h}}\left|\mathcal{P}_{h, h^{\prime}}^{a}(M) \setminus \cup_{h, h^{\prime}}^{a_{1 : t}}\left(m_{t}^{a_{1 : t}}\right)\right| \end{aligned}
+\end{equation*}
+
+Now, for any $a\in \mathcal{J}_h$, let $\tilde{m} = \max_{0\leq t\leq h'} m_t^{a_{1:t}}$. Note that for $m\geq \tilde{m}$, equation \eqref{eq:P-min-size} is not satisfied. Thus we have
+\begin{equation*}
+\begin{aligned}
+\left|\mathcal{P}_{h, h^{\prime}}^{a} \setminus \cup_{h, h^{\prime}}^{s+1} \mathcal{P}_{h, h^{\prime}}^{a_{1 : t}}\left(m_{t}^{a_{1 : t}}\right)\right|=&\sum_{m=\tilde{m}}^{M-1} \tau_{h, h^{\prime}}^{a}(m+1)=\sum_{m=0}^{M-1} \tau_{h, h^{\prime}}^{a}(m+1) \mathbbm{1}\{\eqref{eq:P-min-size} \text { is not satisfied }\} \\ & \leq \sum_{m=0}^{M-1} \tau_{h, h^{\prime}}^{a}(m+1) \mathbbm{1}\{\eqref{eq:cond-ukl} \text { or }\eqref{eq:cond-lkl}\} \end{aligned}
+\end{equation*}
+where the last inequality results from Lemma \ref{lemma:size_Ph}. Hence, we proved:
 
 \begin{equation*}
-    |\mathcal{P}^\emptyset_{h,h'}(M)| \leq \mathbbm{1}_{h'>0}\sum_{t=0}^{h'}(\gamma^{2(t-h')}|\mathcal{I}_t| + \sum_{m=1}^{M}\sum_{a\in \mathcal{J}_h} \mathbbm{1}\{\eqref{eq:cond-ukl}\text{ or }\eqref{eq:cond-lkl}\}
+\left|\mathcal{P}_{h, h^{\prime}}^{\emptyset}\right| \leq \sum_{t=0}^{h^{\prime}} \gamma^{2\left(t-h^{\prime}\right)}\left|\mathcal{I}_{t}\right|+\sum_{m=0}^{M-1} \sum_{a \in \mathcal{J}_{h}} \mathbbm{1}\{\eqref{eq:cond-ukl}\text{ or }\eqref{eq:cond-lkl}\}
 \end{equation*}
-Taking the expectation and applying Lemma \ref{lemma:boundary-crossing-prob} yield the claimed bound.
+
+Taking the expectation and applying Lemma \ref{lemma:boundary-crossing-prob} yield the claimed bound for $h'\geq 1$.
+
+Now for $h' = 0$ we need a modified version of Lemma \ref{lemma:size_Ph}.
+Indeed in this case one can directly prove that $\tau_{h,0}^a(m+1)=1$ implies that either equation \eqref{eq:cond-ukl} or \eqref{eq:cond-lkl} is satisfied (this follows from the fact that $\tau_{h,0}^a(m+1)=1$ always imply that \eqref{eq:sampled-enough-h} is true for $h'= 0$).
+Thus we obtain:
+\begin{equation*}
+\left|\mathcal{P}_{h, h^{\prime}}^{\emptyset}\right|=\sum_{m=0}^{M-1} \sum_{a \in \mathcal{J}_{h}} \tau_{h, 0}^{a}(m+1) \leq \sum_{m=0}^{M-1} \sum_{a \in \mathcal{J}_{h}} \mathbbm{1}\{\eqref{eq:cond-ukl}\text{ or }\eqref{eq:cond-lkl}\}
+\end{equation*}
+Taking the expectation and applying Lemma \ref{lemma:boundary-crossing-prob} yield the claimed bound for $h' = 0$ and ends the proof.
 \qed
@@
 \begin{proof}
-The proof is identical to that of Lemma 10 in \citep{Bubeck2010}, where they show:
-\begin{equation*}
-    \sum_{a\in \mathcal{J}_h} T_a(M) = \tilde{O}\left((\kappa')^h + \gamma^{-2h}\sum_{h'=1}^{h-1} \gamma^{2h'}|\mathcal{P}^\emptyset_{h,h'}(M)| + M|\mathcal{P}^\emptyset_{h,0}(M)|\right)
-\end{equation*}
+The proof is identical to that of Lemma 10 in \citep{Bubeck2010}:
+\begin{align*}
+\sum_{a \in \mathcal{J}_{h}} T_{a}(M) &= \sum_{a \in \mathcal{J}_{h} \backslash \mathcal{P}_{h, h-1}^{\emptyset}} T_{a}(M)+\sum_{h^{\prime}=1}^{h-1} \sum_{a \in \mathcal{P}_{h, h^{\prime}}^{\emptyset} \setminus \mathcal{P}_{h, h^{\prime}-1}^{\emptyset}} T_{a}(M)+\sum_{a \in \mathcal{P}_{h, 0}^{\emptyset}} T_{a}(M)\\
+&\leq 2f(m)(h+1)^{2} \gamma^{2(h-2-h)}\left|\mathcal{J}_{h}\right|+\sum_{h^{\prime}=1}^{h-1} 2f(m)(h+1)^{2} \gamma^{2\left(h^{\prime}-2-h\right)} \log M\left|\mathcal{P}_{h, h^{\prime}}^{\emptyset}\right|+M\left|\mathcal{P}_{h, 0}^{\emptyset}\right|\\
+&=\tilde{O}\left(\left(\kappa^{\prime}\right)^{h}+\gamma^{-2 h} \sum_{h^{\prime}=1}^{h-1} \gamma^{2 h^{\prime}}\left|\mathcal{P}_{h, h^{\prime}}^{\emptyset}\right|+M\left|\mathcal{P}_{h, 0}^{\emptyset}\right|\right)
+\end{align*}
 Taking the expectation and applying the bound of Lemma \ref{lemma:expected-P-size} give the claimed bound.

### 2019-04-13 paper/main.tex
Updates from Overleaf

@@
 In stark contrast, the \emph{Optimistic Planning for Deterministic systems} (\texttt{OPD}) algorithm considered by \citet{Hren2008} in the case of deterministic transitions and rewards exploits the structure of the cumulative discounted reward to achieve a problem-dependent polynomial bound on sample complexity.
-A similar line of work in a deterministic setting is that of \texttt{SOOP} and \texttt{OPC} by \cite{Busoniu2013,Busoniu2018} thoush they focus on continuous action spaces.
+A similar line of work in a deterministic setting is that of \texttt{SOOP} and \texttt{OPC} by \cite{Busoniu2013,Busoniu2018} though they focus on continuous action spaces.
 \texttt{OPD} was later extended to stochastic systems with the \emph{Open-Loop Optimistic Planning} (\OLOP) algorithm introduced by \citet{Bubeck2010} in the open-loop setting: we only consider sequences of actions independently of the states that they lead to.
@@
 Indeed, as long as a subtree under the root verifies $U^{\mu}_a(m) > 1$ for every sequence $a$, then all these sequences share the same B-value $B_a(m) = U_{a_{1:1}(m)}$.
-This means that \OLOP cannot differentiate them and exploit information from their shared history as intended, and behaves as uniform sampling instead.
+This means that \OLOP cannot discriminate them and exploit information from their shared history as intended, and behaves as uniform sampling instead.
 Once the early depths have been explored sufficiently, \OLOP resumes its intended behaviour, but the problem is only shifted to deeper unexplored subtrees.
@@
     \centering
+
+    \includesvg[width=0.64\textwidth]{img/ODP}
     \includesvg[width=0.64\textwidth]{img/OLOP}
     \includesvg[width=0.64\textwidth]{img/KL-OLOP}
-    \includesvg[width=0.64\textwidth]{img/ODP}
     \caption{The look-ahead trees expanded by different planning algorithms for a budget $n=10^3$, shown down to depth 6. The width of edges is proportional to the nodes visit count $T_a(M)$.}

### 2019-04-13 paper/main.tex
Use lmodern font
Arxiv-friendly figures

@@
-\let\origvec\vec
+\let\accentvec\vec            %add before \documentclass
 \documentclass[runningheads]{llncs}
-\let\springervec\vec
-\let\vec\origvec
+\let\spvec\vec                %add before \usepackage{amsmath}
+\let\vec\accentvec
 
@@
 \usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{lmodern}
 \usepackage{url}            % simple URL typesetting
@@
 %\usepackage{natbib}
+\usepackage[cmex10]{amsmath}
+\usepackage{amssymb}
 \usepackage{hyperref}
@@
 \usepackage{float}
-\usepackage{amsmath}
-\usepackage{amssymb}
+
 \usepackage{calc}
@@
 \usepackage{svg}
-\usepackage[lofdepth,lotdepth]{subfig}
-\usepackage{comment}
+\usepackage{subfig}
+%\usepackage{comment}
 
@@
     \centering
-    \includegraphics[width=0.6\textwidth]{img/ukl.eps}
+    \includegraphics[width=0.6\textwidth]{img/ukl-eps-converted-to}
     \caption{The Bernoulli Kullback-Leibler divergence $\dber$, and the corresponding upper and lower confidence bounds $U^{\mu}_a$ and $L^{\mu}_a$ for the empirical average $\hat{\mu_a}$.
@@
     \label{eq:kappa}
-    \kappa_2 = \limsup_{h\rightarrow\infty}{\left|\left\{a\in a^H:V(a) \geq V - 2\frac{\gamma^{h+1}}{1-\gamma}\right\}\right|^{1/h}}
+    \kappa_2 \eqdef \limsup_{h\rightarrow\infty}{\left|\left\{a\in a^H:V(a) \geq V - 2\frac{\gamma^{h+1}}{1-\gamma}\right\}\right|^{1/h}}
 \end{equation*}
@@
     \centering
-    \includesvg[width=0.6\textwidth]{img/tree}
+    \includegraphics[width=0.6\textwidth]{img/tree_svg-tex}
     \caption{A representation of the tree $\Tau_m^+$, with $K = 2$ actions and after episode $m = 2$, when two sequences have been sampled.
@@
 \centering
-\begin{minipage}{\textwidth}
-    \subfloat[Highway]{
-      \includesvg[width=\textwidth]{img/hw_return.svg}
+%\begin{minipage}{\textwidth}
+%      \includegraphics[width=\textwidth]{img/hw_return_svg-tex}
+%      \label{sub:highway}
+%      \includegraphics[width=0.5\textwidth]{img/gw_return_svg-tex}
+%      \label{sub:gridworld}
+%      \includegraphics[width=0.5\textwidth]{img/gw_stoch_return_svg-tex}
+%      \label{sub:gridworld_stoch}
+
+\subfloat{
+	\includegraphics[width=\textwidth]{img/hw_return_svg-tex}
 	\label{sub:highway}
@@
 \newline
-    \subfloat[Gridworld]{
-      \includesvg[width=0.5\textwidth]{img/gw_return.svg}
+\subfloat{
+	\includegraphics[width=0.5\textwidth]{img/gw_return_svg-tex}
 	\label{sub:gridworld}
 }
-    \subfloat[Stochastic gridworld]{
-      \includesvg[width=0.5\textwidth]{img/gw_stoch_return.svg}
+\subfloat{
+	\includegraphics[width=0.5\textwidth]{img/gw_stoch_return_svf-tex}
 	\label{sub:gridworld_stoch}
 }
-
     \caption{Numerical experiments: for each environment-agent configuration, we compute the average return over 100 runs –- along with its 95\% confidence interval –- with respect to the available budget $n$.}
     \label{fig:experiments}
-\end{minipage}
+%\end{minipage}
 \end{figure}
@@
 The results of our experiments are shown in Figure \ref{fig:experiments}.
-The \texttt{ODP} algorithm converges very quickly to the optimal return in the two first environments shown in Figure \ref{sub:highway} and Figure \ref{sub:gridworld}, because it exploits their deterministic nature: it needs neither to estimate the rewards through upper-confidence bounds nor to sample whole sequences all the way from the root when expanding a leaf, which provides a significant speedup.
+The \texttt{ODP} algorithm converges very quickly to the optimal return in the two first environments, shown in Figure \ref{sub:highway} and Figure \ref{sub:gridworld}, because it exploits their deterministic nature: it needs neither to estimate the rewards through upper-confidence bounds nor to sample whole sequences all the way from the root when expanding a leaf, which provides a significant speedup.
 It can be seen of an oracle allowing to measure the conservativeness of stochastic planning algorithms.
@@
 \bibliographystyle{splncs04}
-\bibliography{references}
+\begin{thebibliography}{10}
+\providecommand{\url}[1]{\texttt{#1}}
+\providecommand{\urlprefix}{URL }
+\providecommand{\doi}[1]{https://doi.org/#1}
+
+\bibitem{Bellman1957}
+Bellman, R.: {Dynamic Programming}. Princeton University Press  (1957)
+
+\bibitem{Bubeck2010}
+Bubeck, S., Munos, R.: {Open Loop Optimistic Planning}. In: Proc. of COLT
+  (2010)
+
+\bibitem{Busoniu2013}
+Buşoniu, L., Daniels, A., Munos, R., Babuska, R.: {Optimistic planning for
+  continuous-action deterministic systems}. IEEE Symposium on Adaptive Dynamic
+  Programming and Reinforcement Learning, ADPRL  (2013)
+
+\bibitem{Busoniu2018}
+Buşoniu, L., P{\'{a}}ll, E., Munos, R.: {Continuous-action planning for
+  discounted infinite-horizon nonlinear optimal control with Lipschitz values}.
+  Automatica  \textbf{92}(December),  100--108 (2018)
+
+\bibitem{Cappe2013}
+Capp{\'{e}}, O., Garivier, A., Maillard, O.A., Munos, R., Stoltz, G.:
+  {Kullback-Leibler Upper Confidence Bounds for Optimal Sequential Allocation}.
+  The Annals of Statistics  \textbf{41}(3),  1516--1541 (2013)
+
+\bibitem{gym_minigrid}
+Chevalier-Boisvert, M., Willems, L., Pal, S.: Minimalistic gridworld
+  environment for openai gym. \url{https://github.com/maximecb/gym-minigrid}
+  (2018)
+
+\bibitem{Coquelin2007}
+Coquelin, P.A., Munos, R.: {Bandit Algorithms for Tree Search}. Proc. of UAI
+  (2007)
+
+\bibitem{Coulom2006}
+Coulom, R.: {Efficient Selectivity and Backup Operators in Monte-Carlo Tree
+  Search}. In: {Proc. of International Conference on Computer and Games} (2006)
+
+\bibitem{Grill2016}
+Grill, J.B., Valko, M., Munos, R.: {Blazing the trails before beating the path:
+  Sample-efficient Monte-Carlo planning}. In: Proc. of NeurIPS (2016)
+
+\bibitem{Hren2008}
+Hren, J.F., Munos, R.: {Optimistic planning of deterministic systems}. Lecture
+  Notes in Computer Science  (2008)
+
+\bibitem{Kearns2002}
+Kearns, M., Mansour, Y., Ng, A.Y.: {A sparse sampling algorithm for
+  near-optimal planning in large Markov decision processes}. In: Proc. of IJCAI
+  (2002)
+
+\bibitem{Kocsis2006}
+Kocsis, L., Szepesv\'{a}ri, C.: Bandit based monte-carlo planning. In: Proc. of
+  ECML (2006)
+
+\bibitem{Leurent2018}
+Leurent, E., Blanco, Y., Efimov, D., Maillard, O.A.: {Approximate Robust
+  Control of Uncertain Dynamical Systems}. NeurIPS Machine Learning for
+  Intelligent Transportation Systems Workshop  (2018)
+
+\bibitem{Munos2014}
+Munos, R.: {From Bandits to Monte-Carlo Tree Search: The Optimistic Principle
+  Applied to Optimization and Planning}. Foundations and
+  Trends{\textregistered} in Machine Learning  (2014)
+
+\bibitem{Szorenyi2014}
+Szorenyi, B., Kedenburg, G., Munos, R.: {Optimistic planning in Markov decision
+  processes using a generative model}. In: Proc. of NeurIPS (2014)
+
+\end{thebibliography}
+
 %\begin{comment}
@@
 They were all generated in the same initial root state, under the same budget of $n=10^3$ calls to the generative model.
-We observe that the tree developed by the \OLOP algorithm is quite evenly balanced, which suggests that \OLOP behaves as uniform planning as hypothesised in \ref{sec:kl-olop-behaviour}.
-On the contrary, \KLOLOP is much more efficient and produced a highly unbalanced tree, with most of the sampled sequences densely concentrated around the optimal trajectory, while the other branches under the root have barely been sampled more than once.
-\ODP is even more efficient in this deterministic setting, as it does not need to run the trajectories from the root down to depth $L$, nor to explore twice the same transition.
+We observe that \ODP exploits the deterministic setting efficiently and produces a very sparse tree densely concentrated around the optimal trajectory.
+Conversely, the tree developed by the \OLOP algorithm is quite evenly balanced, which suggests that \OLOP behaves as uniform planning as hypothesised in \ref{sec:kl-olop-behaviour}.
+In stark contrast, \KLOLOP is much more efficient and produced a highly unbalanced tree, exploring the same regions of the tree as \ODP.
 
@@
 
-    \includesvg[width=0.64\textwidth]{img/ODP}
-    \includesvg[width=0.64\textwidth]{img/OLOP}
-    \includesvg[width=0.64\textwidth]{img/KL-OLOP}
+    \includegraphics[width=0.64\textwidth]{img/ODP_svg-tex}    \includegraphics[width=0.64\textwidth]{img/OLOP_svg-tex}
+    \includegraphics[width=0.64\textwidth]{img/KL-OLOP_svg-tex}
+
     \caption{The look-ahead trees expanded by different planning algorithms for a budget $n=10^3$, shown down to depth 6. The width of edges is proportional to the nodes visit count $T_a(M)$.}

### 2019-04-13 paper/main.tex
Fix wording

@@
 Indeed, as long as a subtree under the root verifies $U^{\mu}_a(m) > 1$ for every sequence $a$, then all these sequences share the same B-value $B_a(m) = U_{a_{1:1}(m)}$.
-This means that \OLOP cannot discriminate them and exploit information from their shared history as intended, and behaves as uniform sampling instead.
+This means that \OLOP cannot differentiate them and exploit information from their shared history as intended, and behaves as uniform sampling instead.
 Once the early depths have been explored sufficiently, \OLOP resumes its intended behaviour, but the problem is only shifted to deeper unexplored subtrees.

### 2019-06-11 paper/main.tex
Updates from Overleaf

@@
     \centering
-    \includegraphics[width=0.6\textwidth]{img/ukl-eps-converted-to}
+    \includegraphics[width=0.6\textwidth]{img/ukl}
     \caption{The Bernoulli Kullback-Leibler divergence $\dber$, and the corresponding upper and lower confidence bounds $U^{\mu}_a$ and $L^{\mu}_a$ for the empirical average $\hat{\mu_a}$.
@@
     \item \texttt{KL-OLOP}(1): an aggressive version of \KLOLOP where we used $f_1(m) = \log M$ instead of $f_2(m)$.
-This threshold function makes the upper bounds even tighter, at the cost of an increase probability of violation.
+This threshold function makes the upper bounds even tighter, at the cost of an increased probability of violation.
 Hence, we expect this solution to be more efficient in close-to-deterministic environments.
@@
 The \texttt{ODP} algorithm converges very quickly to the optimal return in the two first environments, shown in Figure \ref{sub:highway} and Figure \ref{sub:gridworld}, because it exploits their deterministic nature: it needs neither to estimate the rewards through upper-confidence bounds nor to sample whole sequences all the way from the root when expanding a leaf, which provides a significant speedup.
-It can be seen of an oracle allowing to measure the conservativeness of stochastic planning algorithms.
+It can be seen as an oracle allowing to measure the conservativeness of stochastic planning algorithms.
 And indeed, even before introducing stochasticity, we can see that \OLOP performs quite badly on the two environments, only managing to solve them with a budget in the order of $10^{3.5}$.

### 2019-06-11 paper/poster/poster.tex
Add poster structure

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% baposter Landscape Poster
% LaTeX Template
% Version 1.0 (15/5/13)
%
% Created by:
% Brian Amberg (baposter@brian-amberg.de)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[paperwidth=36in, paperheight=48in,portrait,fontscale=0.35, margin=2cm]{baposter}

\usepackage[utf8]{inputenc}
\usepackage[font=normalsize,labelfont=bf]{caption} % Required for specifying captions to tables and figures
\captionsetup{justification=centering}
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{lmodern}
\usepackage{relsize} % Used for making text smaller in some places

% Other packages
\usepackage{tabularx}
\def\tabularxcolumn#1{m{#1}}
\def\imagetop#1{\vtop{\null\hbox{#1}}}
\usepackage{adjustbox}
\usepackage{subfig}
\usepackage{floatrow} % sidecapfloat
\usepackage{setspace}
\usepackage{xspace}
\usepackage{noReferences}
\usepackage{caption}
\usepackage{tikz}
\usepackage[customcolors]{hf-tikz}
%\usepackage{subcaption}
\usepackage{epsf}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{dsfont}
\usepackage{enumitem}
\usepackage{float}
\usepackage{cases}
\usepackage{verbatim}
\usepackage{array}
\usepackage{graphicx}
\usepackage{graphics}
\usepackage{multirow}
\usepackage{listings}
%\usepackage[authoryear]{natbib}
\usepackage{authblk}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{eqparbox}
\usepackage{xcolor}
\usepackage[backend=biber, citestyle=authoryear, maxcitenames=2, maxbibnames=2]{biblatex}

\bibliography{../references.bib}

\input{style}
\input{../mathdef}

\begin{document}

%\background{ % Set the background to an image (background.pdf)
%\begin{tikzpicture}[remember picture,overlay]
%\draw (current page.north west)+(-2em,2em) node[anchor=north west]
%{\includegraphics[height=1.1\textheight]{background.png}};
%\end{tikzpicture}
%}
\begin{poster}{
grid=false,
borderColor=bordercol, % Border color of content boxes
headerColorOne=headercol1, % Background color for the header in the content boxes (left side)
headerColorTwo=headercol2, % Background color for the header in the content boxes (right side)
headerFontColor=headerfontcol, % Text color for the header text in the content boxes
boxColorOne=boxcolor, % Background color for the content in the content boxes
headershape=roundedright, % Specify the rounded corner in the content box headers
%headerfont=\Large\sf\bf, % Font modifiers for the text in the content box headers
headerfont=\Large\bf\textsc, %Sans Serif
textborder=rectangle,
background=none,
headerborder=open, % Change to closed for a line under the content box headers
boxshade=plain,
textfont={\setlength{\parindent}{0.0em}\sffamily},
headerheight={0.05\textheight},
eyecatcher=true
%columns=5
}
%
%----------------------------------------------------------------------------------------
%	Title and authors
%----------------------------------------------------------------------------------------
%
{
\includegraphics[width=14em]{./img/renault_group}
}
{
Approximate Robust Control of Uncertain\\Dynamical Systems
}
{
Edouard Leurent, Yann Blanco, Denis Efimov, Odalric-Ambrym Maillard
\vspace{-3\baselineskip}
}
{
\includegraphics[width=14em]{./img/inria_sc}
}

\setlength{\colheight}{0.92\textheight}

%----------------------------------------------------------------------------------------
%	Motivation
%----------------------------------------------------------------------------------------

\headerbox{\textsc{Motivation}}{name=motivation,span=1,column=0,row=0}{
\begin{center}
%\includegraphics[width=\textwidth]{../img/pipeline}    
\textbf{Decision-making} pipeline in Autonomous Driving
\end{center}

\textbf{Progress has been made} {\small \parencite{Paden2016}}
\begin{itemize}[nolistsep]
\item[$\eqgcolb{\blacktriangleright}$] \hlg{route planning} is solved
\item[$\eqbcolb{\blacktriangleright}$] reliable techniques exist for \hlb{motion planning} and \hlb{control}
\end{itemize}

\vspace{.5\baselineskip}
\textbf{Current limitations}
\begin{itemize}[nolistsep,leftmargin=0.55cm]
\item[$\eqrcolb{\blacktriangleright}$] Implementations rely on \hlr{hand-crafted rules} such as FSM
\item[] \incarrow tailored for \hlr{specific} cases, won't \hlr{scale} to complex scenes
\item[$\eqrcolb{\blacktriangleright}$] Social interactions are difficult to model explicitly:
\item[] \incarrow poor \hlr{negotiation} abilities, \hlr{"freezing robot"} problem
\end{itemize}

\vspace{.5\baselineskip}
\textbf{We frame the problem in a sequential learning setting}
}

%----------------------------------------------------------------------------------------
%	Setting
%----------------------------------------------------------------------------------------
\headerbox{\textsc{Reinforcement learning}}{name=setting,span=1,column=0,row=0,below=motivation}{

    \hlb{\textbf{Optimal}} control under \hlr{unknown} dynamics $T(s_{t+1} | s_t, a_t)$

    \begin{equation}
        \max_\pi \expectedvalue \underbrace{\left[\sum_{t=0}^\infty \gamma^t r(s_t, a_t) \condbar \hleg{a_t\sim\pi(s_t)}, \hler{s_{t+1}\sim T(s_t, a_t)}\right]}_{\text{policy return }R^T_\pi}
    \end{equation}

    \textbf{Model-free} methods directly optimize $\hleg{\pi(a_t | s_t)}$ through policy evaluation and policy improvement.
    \vspace{1em}

    \textbf{Model-based} methods
    \begin{enumerate}[itemsep=0pt]
        \item Learn a model for the dynamics $\hler{\hat{T}(s_t, a_t)}$
        \item (Planning) Leverage it to compute
        \begin{equation*}
        \max_\pi \expectedvalue \left[\sum_{t=0}^\infty \gamma^t r(s_t, a_t) \condbar \hleg{a_t\sim\pi(s_t)}, \hler{s_{t+1}\sim \hat{T}(s_t, a_t)}\right]
        \end{equation*}
    \end{enumerate}

    \begin{itemize}[nolistsep]
        \item[$\eqgcolb{\blacktriangleright}$] Better \hlg{sample efficiency}, \hlg{interpretability}
        \item[$\eqrcolb{\blacktriangleright}$] $\hler{\text{Model bias: }\hat{T} \neq T}$
    \end{itemize}

}

%----------------------------------------------------------------------------------------
% Tools
%----------------------------------------------------------------------------------------
\headerbox{\textsc{Robust optimization}}
%{name=tools,column=2,span =2}
{name=tools,column=0,below=setting,span =1}
{
\begin{enumerate}[nolistsep]
    \item Build a \textbf{confidence region} $\bT$ around $\hler{T}$
    \begin{equation*}
    \forall T'\in \bT, \quad \probability(||\hler{T} - T'|| > \epsilon) < \delta
    \end{equation*}
    \item Plan \textbf{robustly} with respect to this ambiguity
    \begin{equation}
        \hleg{\max_{\pi}}\ \underbrace{\hler{\min_{T\in \bT}} \expectedvalue \sum_{t=0}^\infty \gamma^t r_t}_{v^r(\pi)} 
    \end{equation}
\end{enumerate}

\textbf{One-step} game between the planner and the environment:
\begin{itemize}[nolistsep]
    \item[1.] the \hlg{learner} reveals its policy $\hleg{\pi}$
    \item[2.] the \hlr{adversary} chooses the worst-case dynamics $T$
\end{itemize}

\begin{center}
    \hlb{\textbf{Assumption}}
\end{center}
We consider deterministic systems: $s_{t+1} = T(s_t, a_t)$

\begin{center}
\hlr{\textbf{Challenge}} 
\end{center}
How to optimize this objective?
\begin{itemize}[nolistsep]
    \item \hlg{Linear} system: $\mathcal{H}_\infty$ control, robust LQ
    \item \hlg{Finite} state-space: Robust Dynamic Programming
    \item \hlr{Non-linear continuous} system: ?
\end{itemize}

}

%----------------------------------------------------------------------------------------
%	Discrete Ambiguity
%----------------------------------------------------------------------------------------
\headerbox{\textsc{Discrete ambiguity and tree-based planning}}{name=discrete,column=1,span=2}{
\vspace{-.25\baselineskip}

\begin{center}
    \hlb{\textbf{Assumption}}
\end{center}
The ambiguity set $\bT$ and the action space $\mathcal{A}$ are discrete and \textbf{finite}: $
 \hleb{\bT = \{T_m\}_{m\in[1, M]}} \quad \text{and} \quad \mathcal{A} = \{a_k\}_{k\in[1, K]}$

 We propose a robust version of \textbf{optimistic planning} for deterministic systems {\small \parencite{Hren2008}}

\begin{tabular}{>{\centering}p{0.63\textwidth}>{\centering}p{0.32\textwidth}}
\begin{minipage}[]{0.63\textwidth}
\bookboxx{
\begin{definition}
Given node $i\in\mathcal{T}$, define

\textbf{The robust value:}
$$v^r_i \eqdef \hleg{\vphantom{\min_{m\in[1, M]}}\max_{\pi \in i\mathcal{A^\infty}}} \hler{\min_{m\in[1, M]}} R^{T_m}_\pi$$

\textbf{The robust u-value:}
$$u_i^r(n) \eqdef
\begin{cases}
\hler{\min_{m\in[1, M]}} \sum_{t=0}^{d-1} \gamma^t r_t &\text{if } i \in \mathcal{L}_n \text{ ;}\\
\hleg{\max_{a\in\mathcal{A}}}\ u_{ia}^r(n) & \text{if } i \in \mathcal{T}_n \setminus \mathcal{L}_n
\end{cases}$$

\textbf{The robust b-value:}
$$b_i^r(n) \eqdef
\begin{cases}
\hler{\min_{m\in[1, M]}} \sum_{t=0}^{d-1} \gamma^t r_t + \frac{\gamma^d}{1-\gamma} &\text{if } i \in \mathcal{L}_n \text{ ;}\\
\hleg{\max_{a\in\mathcal{A}}}\ b_{ia}^r(n) & \text{if } i \in \mathcal{T}_n \setminus \mathcal{L}_n 
\end{cases}$$
\end{definition}
}

\bookboxx{
\begin{remark}[Ordering of min and max]
Naive comparison of action values between the different models do not recover the robust policy
%\includegraphics[width=\textwidth]{../img/min-max-order.eps}
\end{remark}
}

\bookboxx{
\begin{theorem}[Regret bound]
\label{theorem:drop-regret}
Algorithm \ref{algo:drop} enjoys a simple regret of:
\vspace{-0.5\baselineskip}
\begin{equation}
\text{If } \hler{\vph\kappa}>1,\qquad 
\hleb{\vph\mathcal{R}_n} = O\left(\hleg{\vph n}^{-\frac{\log 1/\gamma}{\log \hler{\vph\kappa}}}\right)
\end{equation}
\vspace{-\baselineskip}
\begin{equation}
\text{If }\hler{\vph\kappa}=1,\qquad
\hleb{\vph\mathcal{R}_n} = O\left(\gamma^{\frac{(1-\gamma)^\beta}{c}\hleg{\vph n}}\right)
\end{equation}
\vspace{-\baselineskip}
\end{theorem}
}
\end{minipage}

&

\begin{minipage}[]{0.32\textwidth}
\begin{center}
%\includegraphics[width=.95\textwidth]{../img/robust-control-tree}    
\end{center}

\textbf{Variables}
\begin{itemize}[nolistsep]
    \item[\incarrow] computational \hlg{budget $n$}
    \item[\incarrow] near-optimal \hlr{branching factor $\kappa$}
    \item[\incarrow] simple regret \hlb{$\mathcal{R}_n = v^r - v^r_{a(n)}$}
\end{itemize}
\vspace{\baselineskip}
\begin{normalsize}
\begin{algorithm}[H]
\DontPrintSemicolon
Initialize $\mathcal{T}$ to a root and expand it. Set $n=1$.\;
\While{Numerical resource available}{
Compute the robust u-values $u^r_i(n)$ and robust b-values $b^r_i(n)$.\;
Expand $\argmax_{i\in \mathcal{L}_n} b^r_i(n)$.\;
n = n + 1\;
}
\Return $a(n) = \argmax_{a\in \mathcal{A}} u^r_a(n)$
\caption{Deterministic Robust Optimistic Planning}
\label{algo:drop}
\end{algorithm}
\end{normalsize}

\end{minipage}
\end{tabular}
}

%----------------------------------------------------------------------------------------
%	Continuous Ambiguity
%----------------------------------------------------------------------------------------
\headerbox{\textsc{Continuous ambiguity and interval-based planning}}{name=continuous,span=2,column=1,below=discrete}{

\begin{tabular}[T]{cc}
\begin{minipage}{0.49\textwidth}
\hlr{\textbf{Approximate}} the robust objective by a \hlg{tractable} surrogate. \vspace{-0.5\baselineskip}
\bookboxx{
\begin{definition} Given a policy $\pi$ and current state $s_0$, define
\textbf{The reachability set} $S$ at time $t$: 
$$S(t, s_0, \pi) \eqdef \{s_t: \exists T \in \bT \text{ s.t. } s_{k+1} = T(s_{k}, \pi(s_{k}))\}$$

\textbf{The interval hull} {\small \parencite{Puig2005}}
$$\underline{s}(t, s_0, \pi) \eqdef \min S(t, s_0, \pi) \qquad \overline{s}(t, s_0, \pi) \eqdef \max S(t, s_0, \pi)$$
\end{definition}

\textbf{The surrogate objective} $\hat{v^r}$
\begin{equation}
\hleb{\vphantom{\frac{A}{A}}\hat{v^r}(\pi)} \eqdef \sum_{t=0}^H \gamma^t \min_{s\in \square S(t, s_0, \pi)}  r(s, \pi(s))
\end{equation}
}

The approximate performance of a policy is \hlg{guaranteed} on the true environment.
\vspace{-0.5\baselineskip}
\bookboxx{
\begin{proposition}[Lower bound]
\label{prop:lower-bound}
The surrogate objective $\hat{v^r}$ is a lower bound of the true objective $v^r$:

\begin{equation}
\forall \pi,\ \hleb{\vphantom{\frac{A}{A}}\hat{v^r}(\pi)} \leq \hler{\vphantom{\frac{A}{A}}v^r(\pi)}
\end{equation}
\end{proposition}
}

\end{minipage}

&

\begin{minipage}{0.47\textwidth}

\centering
%\includegraphics[width=0.7\textwidth]{../img/interval-hull}

\begin{footnotesize}
\begin{algorithm}[H]
  \SetAlgoLined\DontPrintSemicolon
  \SetKwFunction{algo}{robust\_control}\SetKwFunction{proc}{evaluate}
  \SetKwProg{myalg}{Algorithm}{}{}
  \myalg{\algo{$s_0$}}{
  Initialize a set $\Pi$ of policies\;
  \While{resources available}
  {
  \proc{} each policy $\pi\in\Pi$ at current state $s_0$\; 
  Update $\Pi$ by policy search\;  
  }
  \KwRet $\argmax_{\pi\in\Pi} \hleb{\hat{v^r}(\pi)}$ \;}{}
  \setcounter{AlgoLine}{0}
  \SetKwProg{myproc}{Procedure}{}{}
  \myproc{\proc{$\pi$, $s_0$}}{
  \KwRet \hlb{$\hat{v^r}(\pi)$}\;}
\caption{Interval-based Robust Control}
\label{algo:irc}
\end{algorithm}
\end{footnotesize}
\end{minipage}
\end{tabular}

\vspace{.75\baselineskip}
}

%
%%----------------------------------------------------------------------------------------
%% Experiments
%%----------------------------------------------------------------------------------------
%
\headerbox{Experiments}{name=experiments,span=2,column=1,below=continuous, above=bottom}{

We introduce \textsc{highway-env}, a new environment for simulated highway driving and tactical decision-making\footnote{Video and source code are available at https://eleurent.github.io/robust-control/}.

In these experiments, the \hlg{ego-vehicle} is approaching a roundabout with
\hly{flowing traffic}.

We first consider ambiguity with respect to the possible \hlr{destination} of each vehicle (fig \subref*{fig:destination}), and then \wrt their \hlr{driving style} (fig \subref*{fig:driving-style}).

\begin{figure}[H] 
  \centering 
  \begin{tabular}{l}
  \subfloat[Discrete ambiguity]{\adjustbox{raise=-3.6pc}{
  %\includegraphics[width=0.27\textwidth]{../img/he-discrete.png}
  }\label{fig:destination}}\ 
  \subfloat[Continuous ambiguity]{\adjustbox{raise=-3.6pc}{
  %\includegraphics[width=0.27\textwidth]{../img/he-interval.png}
  \label{fig:driving-style}}}\
    \subfloat[Results]{
        \begin{tabular}{c c c c} \toprule
         Ambiguity & Agent & Worst-case & Mean $\pm$ std  \\ \midrule
         None & Oracle & $9.83$ & $10.84 \pm 0.16$ \\ \midrule
        \multirow{2}{*}{Discrete} & Nominal & $\hler{\vph 2.09}$ & $8.85 \pm 3.53$ \\
         & Algorithm \ref{algo:drop} & \textbf{$\hleg{\vph 8.99}$} &  \textbf{$10.78 \pm 0.34$} \\ \midrule
        \multirow{2}{*}{Continuous} & Nominal & $\hler{\vph 1.99}$ & $9.95 \pm 2.38$ \\
         & Algorithm \ref{algo:irc} & \textbf{$\hleg{\vph 7.88}$} & \textbf{$10.73 \pm 0.61$} \\ \bottomrule
         \end{tabular}
      }
  \end{tabular}
\end{figure}
}

%----------------------------------------------------------------------------------------
%	Acknowledgements
%----------------------------------------------------------------------------------------
\headerbox{Acknowledgements}{name=ack,column=0,span=1,below=tools}{
This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR).
}

%----------------------------------------------------------------------------------------
%	References
%----------------------------------------------------------------------------------------
\headerbox{References}{name=refs,column=0,span=1,below=ack, above=bottom}{
    {
    \AtNextBibliography{\footnotesize}
    \setlength{\bibitemsep}{3pt}
    \printbibliography[heading=none]
    }
}

\end{poster}

\end{document}

### 2019-06-11 paper/main.tex
Updates from Overleaf

@@
 
-    \includegraphics[width=0.64\textwidth]{img/ODP_svg-tex}    \includegraphics[width=0.64\textwidth]{img/OLOP_svg-tex}
-    \includegraphics[width=0.64\textwidth]{img/KL-OLOP_svg-tex}
+    \includegraphics[width=0.64\textwidth]{img/tree-ODP_svg-tex}    \includegraphics[width=0.64\textwidth]{img/tree-OLOP_svg-tex}
+    \includegraphics[width=0.64\textwidth]{img/tree-KL-OLOP_svg-tex}
 

### 2019-06-11 paper/poster/poster.tex
Updates from Overleaf

@@
 \usepackage[backend=biber, citestyle=authoryear, maxcitenames=2, maxbibnames=2]{biblatex}
+\usepackage{lipsum}
+\usepackage{svg}
 
@@
 \input{../mathdef}
+
+% Removed from mathdef because of conflict with paper lcns class
+\newtheorem{theorem}{Theorem}
+\newtheorem{definition}{Definition}
+\newtheorem{lemma}{Lemma}
+\newtheorem{proposition}{Proposition}
+\newtheorem{remark}{Remark}
+\newtheorem{property}{Property}
 
@@
 {
-Approximate Robust Control of Uncertain\\Dynamical Systems
+Practical Open-Loop Optimistic Planning
 }
 {
-Edouard Leurent, Yann Blanco, Denis Efimov, Odalric-Ambrym Maillard
+Edouard Leurent, Odalric-Ambrym Maillard
 \vspace{-3\baselineskip}
@@
 \headerbox{\textsc{Motivation}}{name=motivation,span=1,column=0,row=0}{
-\begin{center}
-%\includegraphics[width=\textwidth]{../img/pipeline}    
-\textbf{Decision-making} pipeline in Autonomous Driving
-\end{center}
 
-\textbf{Progress has been made} {\small \parencite{Paden2016}}
-\begin{itemize}[nolistsep]
-\item[$\eqgcolb{\blacktriangleright}$] \hlg{route planning} is solved
-\item[$\eqbcolb{\blacktriangleright}$] reliable techniques exist for \hlb{motion planning} and \hlb{control}
+We consider the optimal control of an MDP $\cM = (\cS, \cA, R, T, \gamma)$ with bounded rewards $R$
+\begin{itemize}
+    \item $R$ and $T$ are \hlr{unknown}
+    \item Access to a \hlb{generative model} $s' \sim \probability{s'|s,a}$ and $r\sim\probability{r|s, a}$
+    \item \hlr{Fixed-budget} setting: the generative model can be queried $n$ times
 \end{itemize}
-
-\vspace{.5\baselineskip}
-\textbf{Current limitations}
-\begin{itemize}[nolistsep,leftmargin=0.55cm]
-\item[$\eqrcolb{\blacktriangleright}$] Implementations rely on \hlr{hand-crafted rules} such as FSM
-\item[] \incarrow tailored for \hlr{specific} cases, won't \hlr{scale} to complex scenes
-\item[$\eqrcolb{\blacktriangleright}$] Social interactions are difficult to model explicitly:
-\item[] \incarrow poor \hlr{negotiation} abilities, \hlr{"freezing robot"} problem
-\end{itemize}
-
-\vspace{.5\baselineskip}
-\textbf{We frame the problem in a sequential learning setting}
 }
@@
 %----------------------------------------------------------------------------------------
-\headerbox{\textsc{Reinforcement learning}}{name=setting,span=1,column=0,row=0,below=motivation}{
+\headerbox{\textsc{Open-Loop Optimistic Planning}}{name=setting,span=1,column=0,row=0,below=motivation}{
+\OLOP algorithm introduced in [\cite{Bubeck2010}].
 
-    \hlb{\textbf{Optimal}} control under \hlr{unknown} dynamics $T(s_{t+1} | s_t, a_t)$
+Split the budget $n$ into $M$ sequences of length $L$.
+%Let $M$ be the largest integer such that $M \log M/(2 \log 1/\gamma) \leq n$\;
+%Let $L = \log M / (2 \log 1/\gamma)$\;
 
-    \begin{equation}
-        \max_\pi \expectedvalue \underbrace{\left[\sum_{t=0}^\infty \gamma^t r(s_t, a_t) \condbar \hleg{a_t\sim\pi(s_t)}, \hler{s_{t+1}\sim T(s_t, a_t)}\right]}_{\text{policy return }R^T_\pi}
-    \end{equation}
+\includesvg[width=\textwidth]{img/olop-explain.svg}
+\includesvg[width=\textwidth]{img/olop-explain-2.svg}
 
-    \textbf{Model-free} methods directly optimize $\hleg{\pi(a_t | s_t)}$ through policy evaluation and policy improvement.
-    \vspace{1em}
-
-    \textbf{Model-based} methods
-    \begin{enumerate}[itemsep=0pt]
-        \item Learn a model for the dynamics $\hler{\hat{T}(s_t, a_t)}$
-        \item (Planning) Leverage it to compute
-        \begin{equation*}
-        \max_\pi \expectedvalue \left[\sum_{t=0}^\infty \gamma^t r(s_t, a_t) \condbar \hleg{a_t\sim\pi(s_t)}, \hler{s_{t+1}\sim \hat{T}(s_t, a_t)}\right]
-        \end{equation*}
-    \end{enumerate}
-
-    \begin{itemize}[nolistsep]
-        \item[$\eqgcolb{\blacktriangleright}$] Better \hlg{sample efficiency}, \hlg{interpretability}
-        \item[$\eqrcolb{\blacktriangleright}$] $\hler{\text{Model bias: }\hat{T} \neq T}$
-    \end{itemize}
+\begin{center}
+\begin{algorithm}[H]
+\DontPrintSemicolon
+\footnotesize
+\For{each episode $m = 1, \cdots, M$}{
+Compute $U_a(m-1)$ from \eqref{eq:Ua} for all $a\in\mathcal{T}$\;
+Compute $B_a(m-1)$ from \eqref{eq:Ba} for all $a\in A^L$\;\label{alg:b_values_compute}
+Sample a sequence with highest B-value: $a^m \in \argmax_{a\in A^L} B_a(m-1)$.\;
+}
+\Return the most played sequence $a(n) \in \argmax_{a\in A^L} T_a(M)$
+\caption{General structure for Open-Loop Optimistic Planning}
+\label{algo:kl-olop}
+\end{algorithm}
+\end{center}
 
@@
 %----------------------------------------------------------------------------------------
-\headerbox{\textsc{Robust optimization}}
-%{name=tools,column=2,span =2}
+\headerbox{\textsc{What's wrong with OLOP?}}
 {name=tools,column=0,below=setting,span =1}
 {
-\begin{enumerate}[nolistsep]
-    \item Build a \textbf{confidence region} $\bT$ around $\hler{T}$
-    \begin{equation*}
-    \forall T'\in \bT, \quad \probability(||\hler{T} - T'|| > \epsilon) < \delta
-    \end{equation*}
-    \item Plan \textbf{robustly} with respect to this ambiguity
+
+Overly pessimistic, especially in the low-budget regime.
+
 \begin{equation}
-        \hleg{\max_{\pi}}\ \underbrace{\hler{\min_{T\in \bT}} \expectedvalue \sum_{t=0}^\infty \gamma^t r_t}_{v^r(\pi)} 
+         U^{\mu}_a(m) = \hat{\mu}_a(m) + \sqrt{\frac{2 \log M}{T_a(m)}}
     \end{equation}
-\end{enumerate}
-
-\textbf{One-step} game between the planner and the environment:
-\begin{itemize}[nolistsep]
-    \item[1.] the \hlg{learner} reveals its policy $\hleg{\pi}$
-    \item[2.] the \hlr{adversary} chooses the worst-case dynamics $T$
+    \begin{equation}
+    \label{eq:Ua}
+        U_a(m) \eqdef \sum_{t = 1}^h \gamma^t U^{\mu}_{a_{1:t}}(m) + \frac{\gamma^{h+1}}{1-\gamma}
+    \end{equation}
+    \begin{equation}
+    \label{eq:Ba}
+        B_a(m) \eqdef \inf_{1 \leq t \leq L} U_{a_{1:t}}(m)
+    \end{equation}
+    Intuitive explanation: 
+    \begin{itemize}
+        \item Unintended behaviour happens when $\hler{U^{\mu}_a(m) > 1}, \forall a$. 
+        \item Then the sequence $(U_{a_{1:t}}(m))_t$ is non-decreasing
+        \item Then \hlr{$B_a(m) = U_{a_{1:1}}(m)$}
     \end{itemize}
 
-\begin{center}
-    \hlb{\textbf{Assumption}}
-\end{center}
-We consider deterministic systems: $s_{t+1} = T(s_t, a_t)$
+\includesvg[width=\textwidth]{img/olop-explain-3.svg}
 
 \begin{center}
-\hlr{\textbf{Challenge}} 
+\texttt{OLOP} behaves as \hlr{uniform planning}!    
 \end{center}
-How to optimize this objective?
-\begin{itemize}[nolistsep]
-    \item \hlg{Linear} system: $\mathcal{H}_\infty$ control, robust LQ
-    \item \hlg{Finite} state-space: Robust Dynamic Programming
-    \item \hlr{Non-linear continuous} system: ?
-\end{itemize}
-
 }
@@
 %----------------------------------------------------------------------------------------
-%	Discrete Ambiguity
+%	KL-OLOP
 %----------------------------------------------------------------------------------------
-\headerbox{\textsc{Discrete ambiguity and tree-based planning}}{name=discrete,column=1,span=2}{
-\vspace{-.25\baselineskip}
+\headerbox{\textsc{Kullback-Leibler OLOP}}{name=klolop,column=1,span=2}{
+
+\begin{minipage}[]{0.59\textwidth}
+ We summon the upper-confidence bound from \texttt{kl-UCB} [\cite{Cappe2013}]:
+    \begin{equation*}
+        U^{\mu}_a(m) \eqdef \max \left\{q\in I: T_a(m) d(\hat{\mu}_a(m), q) \leq f(m) \right\}
+    \end{equation*}
 
     \begin{center}
-    \hlb{\textbf{Assumption}}
+    \begin{tabular}{ccc}
+    \toprule
+        Algorithm & \texttt{OLOP} & \texttt{KL-OLOP} \\
+        \midrule
+        Interval $I$ & $\mathbb{R}$ & [0, 1] \\
+        Divergence $d$ & $d_{\texttt{QUAD}}$ & $d_{\texttt{BER}}$ \\
+        $f(m)$ & $4 \log M$ & $2\log M + 2 \log\log M$\\
+        \bottomrule
+    \end{tabular}
     \end{center}
-The ambiguity set $\bT$ and the action space $\mathcal{A}$ are discrete and \textbf{finite}: $
- \hleb{\bT = \{T_m\}_{m\in[1, M]}} \quad \text{and} \quad \mathcal{A} = \{a_k\}_{k\in[1, K]}$
+    with
+    \begin{align*}
+    d_{\texttt{QUAD}}(p,q) &\eqdef 2(p-q)^2\\
+    d_{\texttt{BER}}(p, q) &\eqdef p \log \frac{p}{q} + (1-p)\log\frac{1-p}{1-q}
+    \end{align*}
+\end{minipage}
+\begin{minipage}[]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{../img/ukl}
+\end{minipage}
 
- We propose a robust version of \textbf{optimistic planning} for deterministic systems {\small \parencite{Hren2008}}
-
-\begin{tabular}{>{\centering}p{0.63\textwidth}>{\centering}p{0.32\textwidth}}
-\begin{minipage}[]{0.63\textwidth}
-\bookboxx{
-\begin{definition}
-Given node $i\in\mathcal{T}$, define
-
-\textbf{The robust value:}
-$$v^r_i \eqdef \hleg{\vphantom{\min_{m\in[1, M]}}\max_{\pi \in i\mathcal{A^\infty}}} \hler{\min_{m\in[1, M]}} R^{T_m}_\pi$$
-
-\textbf{The robust u-value:}
-$$u_i^r(n) \eqdef
-\begin{cases}
-\hler{\min_{m\in[1, M]}} \sum_{t=0}^{d-1} \gamma^t r_t &\text{if } i \in \mathcal{L}_n \text{ ;}\\
-\hleg{\max_{a\in\mathcal{A}}}\ u_{ia}^r(n) & \text{if } i \in \mathcal{T}_n \setminus \mathcal{L}_n
-\end{cases}$$
-
-\textbf{The robust b-value:}
-$$b_i^r(n) \eqdef
-\begin{cases}
-\hler{\min_{m\in[1, M]}} \sum_{t=0}^{d-1} \gamma^t r_t + \frac{\gamma^d}{1-\gamma} &\text{if } i \in \mathcal{L}_n \text{ ;}\\
-\hleg{\max_{a\in\mathcal{A}}}\ b_{ia}^r(n) & \text{if } i \in \mathcal{T}_n \setminus \mathcal{L}_n 
-\end{cases}$$
-\end{definition}
+Conversely,
+\begin{itemize}
+    \item \hlg{$U^{\mu}_a(m) \in I = [0, 1], \forall a$}. 
+    \item The sequence $(U_{a_{1:t}}(m))_t$ is \hlg{non-increasing}
+    \item $B_a(m) = U_a(m)$, the \emph{bound sharpening} step is \hlb{superfluous}.
+\end{itemize}
 }
 
-\bookboxx{
-\begin{remark}[Ordering of min and max]
-Naive comparison of action values between the different models do not recover the robust policy
-%\includegraphics[width=\textwidth]{../img/min-max-order.eps}
-\end{remark}
+%----------------------------------------------------------------------------------------
+%	Sample Complexity
+%----------------------------------------------------------------------------------------
+\headerbox{Sample complexity}{name=sample,span=2,column=1,below=klolop}{ 
+\begin{theorem}[Sample complexity]
+\label{thm:regret}
+\KLOLOP enjoys the same asymptotic regret bounds as \OLOP. More precisely, \KLOLOP satisfies:
+\begin{equation*}
+    \expectedvalue r_n = \begin{cases}
+      \tilde{0}\left(n^{-\frac{\log 1/\gamma}{\log \kappa'}}\right), & \text{if}\ \gamma\sqrt{\kappa'} > 1 \\
+      \tilde{0}\left(n^{-\frac{1}{2}}\right), & \text{if}\ \gamma\sqrt{\kappa'} \leq 1
+    \end{cases}
+\end{equation*}
+\end{theorem}
 }
 
-\bookboxx{
-\begin{theorem}[Regret bound]
-\label{theorem:drop-regret}
-Algorithm \ref{algo:drop} enjoys a simple regret of:
-\vspace{-0.5\baselineskip}
-\begin{equation}
-\text{If } \hler{\vph\kappa}>1,\qquad 
-\hleb{\vph\mathcal{R}_n} = O\left(\hleg{\vph n}^{-\frac{\log 1/\gamma}{\log \hler{\vph\kappa}}}\right)
-\end{equation}
-\vspace{-\baselineskip}
-\begin{equation}
-\text{If }\hler{\vph\kappa}=1,\qquad
-\hleb{\vph\mathcal{R}_n} = O\left(\gamma^{\frac{(1-\gamma)^\beta}{c}\hleg{\vph n}}\right)
-\end{equation}
-\vspace{-\baselineskip}
+%----------------------------------------------------------------------------------------
+%	Time and Memory Complexity
+%----------------------------------------------------------------------------------------
+\headerbox{Time and memory complexity}{name=time,span=2,column=1,below=sample}{
+
+\begin{tabular}{>{\centering}p{0.45\textwidth}>{\centering}p{0.5\textwidth}}
+\begin{minipage}[]{0.45\textwidth}
+
+\textbf{Original \KLOLOP}
+
+\emph{Compute $B_a(m-1)$ from \eqref{eq:Ba} \hlr{for all $a\in A^L$}}
+
+\textbf{Lazy \KLOLOP}
+\begin{center}
+\includegraphics[width=0.8\textwidth]{../img/tree_svg-tex}    
+\end{center}
+
+\begin{theorem}[Consistency]
+\label{thm:consistency}
+Algorithm \ref{algo:lazy-kl-olop} is \hlg{identical} to Algorithm \ref{algo:kl-olop}.
 \end{theorem}
-}
 \end{minipage}
@@
 
-\begin{minipage}[]{0.32\textwidth}
-\begin{center}
-%\includegraphics[width=.95\textwidth]{../img/robust-control-tree}    
-\end{center}
-
-\textbf{Variables}
-\begin{itemize}[nolistsep]
-    \item[\incarrow] computational \hlg{budget $n$}
-    \item[\incarrow] near-optimal \hlr{branching factor $\kappa$}
-    \item[\incarrow] simple regret \hlb{$\mathcal{R}_n = v^r - v^r_{a(n)}$}
-\end{itemize}
-\vspace{\baselineskip}
-\begin{normalsize}
+\begin{minipage}{0.5\textwidth}
     \begin{algorithm}[H]
 \DontPrintSemicolon
-Initialize $\mathcal{T}$ to a root and expand it. Set $n=1$.\;
-\While{Numerical resource available}{
-Compute the robust u-values $u^r_i(n)$ and robust b-values $b^r_i(n)$.\;
-Expand $\argmax_{i\in \mathcal{L}_n} b^r_i(n)$.\;
-n = n + 1\;
+\footnotesize
+Let $\Tau_0^+ = \LL_0^+ = \{\emptyset\}$\;
+\For{each episode $m = 1, \cdots, M$}{
+Compute $U_a(m-1)$ from \eqref{eq:Ua} for all \hlg{$a\in\Tau_{m-1}^+$}\;
+Compute $B_a(m-1)$ from \eqref{eq:Ba} for all \hlg{$a\in \LL_{m-1}^+$}\;
+Sample a sequence with highest B-value: $a \in \argmax_{\hleg{a\in \LL_{m-1}^+}} B_a(m-1)$\;
+Choose an arbitrary continuation $a^m \in aA^{L-|a|}$\tcp*[h]{e.g. uniformly}
+Let $\Tau_m^+ = \Tau_{m-1}^+$ and $\LL_m^+ = \LL_{m-1}^+$\;
+\For{$t=1, \cdots, L$}{
+    \If{$a^m_{1:t} \not \in \Tau_{m}^+$}{
+    \hlb{Add $a^m_{1:t-1}A$ to $\Tau_{m}^+$ and $\LL_{m}^+$}\;
+    \hlb{Remove $a^m_{1:t-1}$ from $\LL_{m}^+$}
     }
-\Return $a(n) = \argmax_{a\in \mathcal{A}} u^r_a(n)$
-\caption{Deterministic Robust Optimistic Planning}
-\label{algo:drop}
+}
+}
+\Return the most played $a(n) \in \argmax_{a\in \LL_m^+} T_a(M)$
+\caption{Lazy Open Loop Optimistic Planning}
+\label{algo:lazy-kl-olop}
 \end{algorithm}
-\end{normalsize}
-
-\end{minipage}
-\end{tabular}
-}
-
-%----------------------------------------------------------------------------------------
-%	Continuous Ambiguity
-%----------------------------------------------------------------------------------------
-\headerbox{\textsc{Continuous ambiguity and interval-based planning}}{name=continuous,span=2,column=1,below=discrete}{
-
-\begin{tabular}[T]{cc}
-\begin{minipage}{0.49\textwidth}
-\hlr{\textbf{Approximate}} the robust objective by a \hlg{tractable} surrogate. \vspace{-0.5\baselineskip}
-\bookboxx{
-\begin{definition} Given a policy $\pi$ and current state $s_0$, define
-\textbf{The reachability set} $S$ at time $t$: 
-$$S(t, s_0, \pi) \eqdef \{s_t: \exists T \in \bT \text{ s.t. } s_{k+1} = T(s_{k}, \pi(s_{k}))\}$$
-
-\textbf{The interval hull} {\small \parencite{Puig2005}}
-$$\underline{s}(t, s_0, \pi) \eqdef \min S(t, s_0, \pi) \qquad \overline{s}(t, s_0, \pi) \eqdef \max S(t, s_0, \pi)$$
-\end{definition}
-
-\textbf{The surrogate objective} $\hat{v^r}$
-\begin{equation}
-\hleb{\vphantom{\frac{A}{A}}\hat{v^r}(\pi)} \eqdef \sum_{t=0}^H \gamma^t \min_{s\in \square S(t, s_0, \pi)}  r(s, \pi(s))
-\end{equation}
-}
-
-The approximate performance of a policy is \hlg{guaranteed} on the true environment.
-\vspace{-0.5\baselineskip}
-\bookboxx{
-\begin{proposition}[Lower bound]
-\label{prop:lower-bound}
-The surrogate objective $\hat{v^r}$ is a lower bound of the true objective $v^r$:
-
-\begin{equation}
-\forall \pi,\ \hleb{\vphantom{\frac{A}{A}}\hat{v^r}(\pi)} \leq \hler{\vphantom{\frac{A}{A}}v^r(\pi)}
-\end{equation}
-\end{proposition}
-}
-
-\end{minipage}
-
-&
-
-\begin{minipage}{0.47\textwidth}
-
-\centering
-%\includegraphics[width=0.7\textwidth]{../img/interval-hull}
-
-\begin{footnotesize}
-\begin{algorithm}[H]
-  \SetAlgoLined\DontPrintSemicolon
-  \SetKwFunction{algo}{robust\_control}\SetKwFunction{proc}{evaluate}
-  \SetKwProg{myalg}{Algorithm}{}{}
-  \myalg{\algo{$s_0$}}{
-  Initialize a set $\Pi$ of policies\;
-  \While{resources available}
-  {
-  \proc{} each policy $\pi\in\Pi$ at current state $s_0$\; 
-  Update $\Pi$ by policy search\;  
-  }
-  \KwRet $\argmax_{\pi\in\Pi} \hleb{\hat{v^r}(\pi)}$ \;}{}
-  \setcounter{AlgoLine}{0}
-  \SetKwProg{myproc}{Procedure}{}{}
-  \myproc{\proc{$\pi$, $s_0$}}{
-  \KwRet \hlb{$\hat{v^r}(\pi)$}\;}
-\caption{Interval-based Robust Control}
-\label{algo:irc}
-\end{algorithm}
-\end{footnotesize}
 \end{minipage}
@@
 
-\vspace{.75\baselineskip}
+\begin{property}[Time and memory complexity]
+\begin{equation*}
+    \frac{C(\texttt{Lazy KL-OLOP})}{C(\texttt{KL-OLOP})} = \frac{\hleg{nK}}{\hler{K^{L}}}
+\end{equation*}
+\end{property}
+
 }
@@
 %
-\headerbox{Experiments}{name=experiments,span=2,column=1,below=continuous, above=bottom}{
+\headerbox{Experiments}{name=experiments,span=2,column=1,below=time, above=bottom}{
 
-We introduce \textsc{highway-env}, a new environment for simulated highway driving and tactical decision-making\footnote{Video and source code are available at https://eleurent.github.io/robust-control/}.
+\textbf{Expanded trees} for a budget $n=10^3$
 
-In these experiments, the \hlg{ego-vehicle} is approaching a roundabout with
-\hly{flowing traffic}.
+    \includesvg[width=0.33\textwidth]{../img/tree-ODP}    \includesvg[width=0.33\textwidth]{../img/tree-OLOP}
+    \includesvg[width=0.33\textwidth]{../img/tree-KL-OLOP}
 
-We first consider ambiguity with respect to the possible \hlr{destination} of each vehicle (fig \subref*{fig:destination}), and then \wrt their \hlr{driving style} (fig \subref*{fig:driving-style}).
+\textbf{Average return} over 100 runs –- along with its 95\% confidence interval –- with respect to the available budget $n$
 
-\begin{figure}[H] 
-  \centering 
-  \begin{tabular}{l}
-  \subfloat[Discrete ambiguity]{\adjustbox{raise=-3.6pc}{
-  %\includegraphics[width=0.27\textwidth]{../img/he-discrete.png}
-  }\label{fig:destination}}\ 
-  \subfloat[Continuous ambiguity]{\adjustbox{raise=-3.6pc}{
-  %\includegraphics[width=0.27\textwidth]{../img/he-interval.png}
-  \label{fig:driving-style}}}\
-    \subfloat[Results]{
-        \begin{tabular}{c c c c} \toprule
-         Ambiguity & Agent & Worst-case & Mean $\pm$ std  \\ \midrule
-         None & Oracle & $9.83$ & $10.84 \pm 0.16$ \\ \midrule
-        \multirow{2}{*}{Discrete} & Nominal & $\hler{\vph 2.09}$ & $8.85 \pm 3.53$ \\
-         & Algorithm \ref{algo:drop} & \textbf{$\hleg{\vph 8.99}$} &  \textbf{$10.78 \pm 0.34$} \\ \midrule
-        \multirow{2}{*}{Continuous} & Nominal & $\hler{\vph 1.99}$ & $9.95 \pm 2.38$ \\
-         & Algorithm \ref{algo:irc} & \textbf{$\hleg{\vph 7.88}$} & \textbf{$10.73 \pm 0.61$} \\ \bottomrule
-         \end{tabular}
-      }
-  \end{tabular}
-\end{figure}
+     \includesvg[width=0.33\textwidth]{../img/hw_return}
+     \includesvg[width=0.33\textwidth]{../img/gw_return}
+     \includesvg[width=0.33\textwidth]{../img/gw_stoch_return}
 }
@@
 %----------------------------------------------------------------------------------------
-\headerbox{Acknowledgements}{name=ack,column=0,span=1,below=tools}{
-This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR).
-}
+% \headerbox{Acknowledgements}{name=ack,column=0,span=1,below=tools}{
+% This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR).
+% }
 
@@
 %----------------------------------------------------------------------------------------
-\headerbox{References}{name=refs,column=0,span=1,below=ack, above=bottom}{
+\headerbox{References}{name=refs,column=0,span=1,below=tools, above=bottom}{
     {

### 2019-06-11 paper/poster/poster.tex
Updates from Overleaf

@@
 
-We consider the optimal control of an MDP $\cM = (\cS, \cA, R, T, \gamma)$ with bounded rewards $R$
+We consider the optimal control of an MDP $\cM = (\cS, cA, R, T, \gamma)$ with bounded rewards $R$
 \begin{itemize}
@@
 \begin{equation*}
-    \frac{C(\texttt{Lazy KL-OLOP})}{C(\texttt{KL-OLOP})} = \frac{\hleg{nK}}{\hler{K^{L}}}
+    \frac{C(\texttt{Lazy KL-OLOP})}{C(\texttt{KL-OLOP})} = \frac{\hleg{nA}}{\hler{A^{L}}}
 \end{equation*}

### 2019-06-24 paper/main.tex
Updates from Overleaf

@@
 \let\accentvec\vec            %add before \documentclass
-\documentclass[runningheads]{llncs}
+\documentclass[runningheads, envcountsame, a4paper]{llncs}
 \let\spvec\vec                %add before \usepackage{amsmath}
@@
 \usepackage{lmodern}
-\usepackage{url}            % simple URL typesetting
+\usepackage[hyphens]{url}            % simple URL typesetting
 \usepackage{booktabs}       % professional-quality tables
@@
 \usepackage{microtype}      % microtypography
+\usepackage[misc]{ifsym}
 
@@
 \title{Practical Open-Loop Optimistic Planning}
+\toctitle{Practical Open-Loop Optimistic Planning}
 %
@@
 %
-\author{Edouard Leurent\inst{1,2}\and
+\author{Edouard Leurent\inst{1,2}(\Letter)\and
 Odalric-Ambrym Maillard\inst{1}}
+\tocauthor{Edouard~Leurent and
+Odalric-Ambrym~Maillard}
 %
-\authorrunning{E. Leurent, O-A. Maillard}
+\authorrunning{E. Leurent and O-A. Maillard}
 % First names are abbreviated in the running head.
@@
 \institute{SequeL team, INRIA Lille - Nord Europe, France\\
-\email{\{edouard.leurent,odalric.maillard\}@inria.fr}\and
-Renault Group, France\\
-\email{edouard.leurent@renault.com}}
+\email{$\{$edouard.leurent,odalric.maillard$\}$@inria.fr}\and
+Renault Group, France
+%\\
+%\email{edouard.leurent@renault.com}
+}
 %
@@
 They were later reformulated in the setting of Multi-Armed Bandits by \citet{Kocsis2006} with their \emph{Upper Confidence bounds applied to Trees} (\texttt{UCT}) algorithm.
-Despite its popularity, \texttt{UCT} has been shown to suffer from several limitations: its sample complexity can be at least doubly-exponential for some problems (e.g.
+Despite its popularity \citep{Silver2016,Silver2017,Silver2018}, \texttt{UCT} has been shown to suffer from several limitations: its sample complexity can be at least doubly-exponential for some problems (e.g.
 when a narrow optimal path is hidden in a suboptimal branch), which is much worse than uniform planning \citep{Coquelin2007}.
@@
 \begin{proof}
-At episode $m = 1, \cdots, M$, we compute and store in memory of the reward upper-bounds and U-values of all nodes in the subtree $\Tau_m^+$.
+At episode $m = 1, \cdots, M$, we compute and store in memory the reward upper-bounds and U-values of all nodes in the subtree $\Tau_m^+$.
 Moreover, the tree $\Tau_m^+$ is constructed iteratively by adding K nodes at most L times at each episode from 0 to $m$.
@@
 By convention, $\mathcal{I}_0 = \{\emptyset\}$. From the definition of $\kappa_2$, we have that for any $\kappa'>\kappa_2$, there exists a constant C such that for any $h \geq 1$,
-\begin{equation*}
-    |\mathcal{I}_h| \leq C {\kappa'}^h
-\end{equation*}
+%\begin{equation*}
+    $|\mathcal{I}_h| \leq C {\kappa'}^h$
+%\end{equation*}
 Hence, we also have $|\mathcal{J}_h| \leq K|\mathcal{I}_{h-1}| = O({\kappa'}^h)$.
@@
 
-In order to bound this quantity, \diff{we reduce the question to the application of a deviation inequality}.
+\diff{In order to bound this quantity, we reduce the question to the application of a deviation inequality.
 For all $1\leq t\leq L$, we have on the event $\{U^{\mu}_{a^*_{1:t}}(m) \leq \mu(a^*_{1:t})\}$ that $\hat{\mu}_{a^*_{1:t}}(m) \leq U^{\mu}_{a^*_{1:t}}(m) \leq \mu(a^*_{1:t}) < 1$.
-Therefore, for all $0 < \delta < 1 - \mu(a^*_{1:t})$, by definition of $U^{\mu}_{a^*_{1:t}}(m)$:
-
-\begin{equation*}
-    d(\hat{\mu}_{a^*_{1:t}}(m), U^{\mu}_{a^*_{1:t}}(m)+\delta) > \frac{f(m)}{T_{a^*_{1:t}}(m)}
-\end{equation*}
-
-As $d$ is continuous on $(0,1)\times[0, 1]$, we have by letting $\delta \leftarrow 0$ that:
-
-\begin{equation*}
-    d(\hat{\mu}_{a^*_{1:t}}(m), U^{\mu}_{a^*_{1:t}}(m)) \geq \frac{f(m)}{T_{a^*_{1:t}}(m)}
-\end{equation*}
-
-Since d is non-decreasing on $[\hat{\mu}_{a^*_{1:t}}(m), \mu(a^*_{1:t})]$,
-
-\begin{equation*}
-    d(\hat{\mu}_{a^*_{1:t}}(m), \mu(a^*_{1:t})) \geq d(\hat{\mu}_{a^*_{1:t}}(m), U^{\mu}_{a^*_{1:t}}(m)) \geq \frac{f(m)}{T_{a^*_{1:t}}(m)}
-\end{equation*}
-
-We have thus shown the following inclusion:
-
-\begin{equation*}
-    \{U^{\mu}_{a^*_{1:t}}(m) \leq \mu(a^*_{1:t})\} \subseteq \left\{ \mu(a^*_{1:t}) > \hat{\mu}_{a^*_{1:t}}(m) \text{ and } d(\hat{\mu}_{a^*_{1:t}}(m), \mu(a^*_{1:t})) \geq \frac{f(m)}{T_{a^*_{1:t}}(m)} \right\}
-\end{equation*}
-
-Decomposing according to the values of $T_{a^*_{1:t}}(m)$ yields:
-
-\begin{equation*}
-    \{U^{\mu}_{a^*_{1:t}}(m) \leq \mu(a^*_{1:t})\} \subseteq \bigcup_{n=1}^m \left\{ \mu(a^*_{1:t}) > \hat{\mu}_{a^*_{1:t}, n} \text{ and } d(\hat{\mu}_{a^*_{1:t}, n}, \mu(a^*_{1:t})) \geq \frac{f(m)}{n} \right\}
-\end{equation*}
-
-\diff{We now apply the deviation inequality} provided in Lemma 2 of Appendix A in \citep{Cappe2013}: $\forall \epsilon > 1$, provided that $0 < \mu(a^*_{1:t}) < 1$,
-
-\begin{equation*}
-\probability{\bigcup_{n=1}^m \left\{ \mu(a^*_{1:t}) > \hat{\mu}_{a^*_{1:t}, n} \text{ and } n \dber(\hat{\mu}_{a^*_{1:t}, n}, \mu(a^*_{1:t})) \geq \epsilon \right\}} \leq e\ceil{\epsilon \log m}e^{-\epsilon}\,.
-\end{equation*}
-
-By choosing $\epsilon = f(m)$, it comes
-\begin{align*}
-    \probability{\eqref{eq:cond-ukl}} &\leq \sum_{t=1}^L \diff{e\ceil{f(m)\log m}e^{-f(m)}} = \diff{O(L\delta_m)}
+Therefore, for all $0 < \delta < 1 - \mu(a^*_{1:t})$, by definition of $U^{\mu}_{a^*_{1:t}}(m)$:}
+
+\begin{equation*}
+    \diff{d(\hat{\mu}_{a^*_{1:t}}(m), U^{\mu}_{a^*_{1:t}}(m)+\delta) > \frac{f(m)}{T_{a^*_{1:t}}(m)}}
+\end{equation*}
+
+\diff{As $d$ is continuous on $(0,1)\times[0, 1]$, we have by letting $\delta \rightarrow 0$ that:}
+
+\begin{equation*}
+    \diff{d(\hat{\mu}_{a^*_{1:t}}(m), U^{\mu}_{a^*_{1:t}}(m)) \geq \frac{f(m)}{T_{a^*_{1:t}}(m)}}
+\end{equation*}
+
+\diff{Since d is non-decreasing on $[\hat{\mu}_{a^*_{1:t}}(m), \mu(a^*_{1:t})]$,}
+
+\begin{equation*}
+    \diff{d(\hat{\mu}_{a^*_{1:t}}(m), \mu(a^*_{1:t})) \geq d(\hat{\mu}_{a^*_{1:t}}(m), U^{\mu}_{a^*_{1:t}}(m)) \geq \frac{f(m)}{T_{a^*_{1:t}}(m)}}
+\end{equation*}
+
+\diff{We have thus shown the following inclusion:}
+
+\begin{equation*}
+    \diff{\{U^{\mu}_{a^*_{1:t}}(m) \leq \mu(a^*_{1:t})\} \subseteq \left\{ \mu(a^*_{1:t}) > \hat{\mu}_{a^*_{1:t}}(m) \text{ and } d(\hat{\mu}_{a^*_{1:t}}(m), \mu(a^*_{1:t})) \geq \frac{f(m)}{T_{a^*_{1:t}}(m)} \right\}}
+\end{equation*}
+
+\diff{Decomposing according to the values of $T_{a^*_{1:t}}(m)$ yields:}
+
+\begin{equation*}
+    \diff{\{U^{\mu}_{a^*_{1:t}}(m) \leq \mu(a^*_{1:t})\} \subseteq \bigcup_{n=1}^m \left\{ \mu(a^*_{1:t}) > \hat{\mu}_{a^*_{1:t}, n} \text{ and } d(\hat{\mu}_{a^*_{1:t}, n}, \mu(a^*_{1:t})) \geq \frac{f(m)}{n} \right\}}
+\end{equation*}
+
+\diff{We now apply the deviation inequality provided in Lemma 2 of Appendix A in \citep{Cappe2013}: $\forall \epsilon > 1$, provided that $0 < \mu(a^*_{1:t}) < 1$,}
+
+\begin{equation*}
+\diff{\probability{\bigcup_{n=1}^m \left\{ \mu(a^*_{1:t}) > \hat{\mu}_{a^*_{1:t}, n} \text{ and } n \dber(\hat{\mu}_{a^*_{1:t}, n}, \mu(a^*_{1:t})) \geq \epsilon \right\}} \leq e\ceil{\epsilon \log m}e^{-\epsilon}\,.}
+\end{equation*}
+
+\diff{By choosing $\epsilon = f(m)$, it comes}
+\begin{align*}
+    \diff{\probability{\eqref{eq:cond-ukl}} \leq \sum_{t=1}^L e\ceil{f(m)\log m}e^{-f(m)} = O(L\delta_m)}
 \end{align*}
@@
 \begin{proof}
-We start by providing an explicit upper-bound for the length of the confidence interval $U^{\mu}_{a_{1:t}} - L^{\mu}_{a_{1:t}}$. By Pinsker's inequality:
+We start by providing an explicit upper-bound for the length of the confidence interval $U^{\mu}_{a_{1:t}} - L^{\mu}_{a_{1:t}}$. \diff{By Pinsker's inequality:}
 
@@
 
-Hence for all $C>0$, 
-\begin{equation*}
-    \dber(p, q) \leq C   \implies 2(q - p)^2 < C  \implies p - \sqrt{C/2} < q < p + \sqrt{C/2}
-\end{equation*}
-And thus, for all $b\in A^*$, by definition of $U^{\mu}$ and $L^{\mu}$:
-\begin{align*}
-    U^{\mu}_{b}(m) - L^{\mu}_{b}(m) &\leq \frac{S_b(m)}{T_b(m)} + \sqrt{\frac{f(m)}{2T_b(m)}} -  \left(\frac{S_b(m)}{T_b(m)} - \sqrt{\frac{f(m)}{2T_b(m)}}\right) 
-    = \diff{\sqrt{\frac{2f(m)}{T_b(m)}}}
+\diff{Hence for all $C>0$, }
+\begin{equation*}
+    \diff{\dber(p, q) \leq C   \implies 2(q - p)^2 < C  \implies p - \sqrt{C/2} < q < p + \sqrt{C/2}}
+\end{equation*}
+\diff{And thus, for all $b\in A^*$, by definition of $U^{\mu}$ and $L^{\mu}$:}
+\begin{align*}
+    \diff{U^{\mu}_{b}(m) - L^{\mu}_{b}(m) \leq \frac{S_b(m)}{T_b(m)} + \sqrt{\frac{f(m)}{2T_b(m)}} -  \left(\frac{S_b(m)}{T_b(m)} - \sqrt{\frac{f(m)}{2T_b(m)}}\right) 
+    = \sqrt{\frac{2f(m)}{T_b(m)}}}
 \end{align*}
@@
 \begin{itemize}
-    \item A \href{https://github.com/eleurent/highway-env/}{highway driving} environment \citep{Leurent2018}: a vehicle is driving on a road randomly populated with other slower drivers, and must make their way as fast as possible while avoiding collisions by choosing on the the following actions: \texttt{change-lane-left}, \texttt{change-lane-right}, \texttt{no-op}, \texttt{faster}, \texttt{slower}.
+    \item A \href{https://github.com/eleurent/highway-env/}{highway driving} environment \citep{highway-env}: a vehicle is driving on a road randomly populated with other slower drivers, and must make their way as fast as possible while avoiding collisions by choosing on the the following actions: \texttt{change-lane-left}, \texttt{change-lane-right}, \texttt{no-op}, \texttt{faster}, \texttt{slower}.
     \item A \href{https://github.com/maximecb/gym-minigrid}{gridworld} environment \citep{gym_minigrid}: the agent navigates in a randomly-generated gridworld composed of either empty cells, terminal lava cells, and goal cells where a reward of $1$ is collected at the first visit.
@@
 
+\section*{Acknowledgments}
+This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche.
+
 \bibliographystyle{splncs04}
-\begin{thebibliography}{10}
-\providecommand{\url}[1]{\texttt{#1}}
-\providecommand{\urlprefix}{URL }
-\providecommand{\doi}[1]{https://doi.org/#1}
-
-\bibitem{Bellman1957}
-Bellman, R.: {Dynamic Programming}. Princeton University Press  (1957)
-
-\bibitem{Bubeck2010}
-Bubeck, S., Munos, R.: {Open Loop Optimistic Planning}. In: Proc. of COLT
-  (2010)
-
-\bibitem{Busoniu2013}
-Buşoniu, L., Daniels, A., Munos, R., Babuska, R.: {Optimistic planning for
-  continuous-action deterministic systems}. IEEE Symposium on Adaptive Dynamic
-  Programming and Reinforcement Learning, ADPRL  (2013)
-
-\bibitem{Busoniu2018}
-Buşoniu, L., P{\'{a}}ll, E., Munos, R.: {Continuous-action planning for
-  discounted infinite-horizon nonlinear optimal control with Lipschitz values}.
-  Automatica  \textbf{92}(December),  100--108 (2018)
-
-\bibitem{Cappe2013}
-Capp{\'{e}}, O., Garivier, A., Maillard, O.A., Munos, R., Stoltz, G.:
-  {Kullback-Leibler Upper Confidence Bounds for Optimal Sequential Allocation}.
-  The Annals of Statistics  \textbf{41}(3),  1516--1541 (2013)
-
-\bibitem{gym_minigrid}
-Chevalier-Boisvert, M., Willems, L., Pal, S.: Minimalistic gridworld
-  environment for openai gym. \url{https://github.com/maximecb/gym-minigrid}
-  (2018)
-
-\bibitem{Coquelin2007}
-Coquelin, P.A., Munos, R.: {Bandit Algorithms for Tree Search}. Proc. of UAI
-  (2007)
-
-\bibitem{Coulom2006}
-Coulom, R.: {Efficient Selectivity and Backup Operators in Monte-Carlo Tree
-  Search}. In: {Proc. of International Conference on Computer and Games} (2006)
-
-\bibitem{Grill2016}
-Grill, J.B., Valko, M., Munos, R.: {Blazing the trails before beating the path:
-  Sample-efficient Monte-Carlo planning}. In: Proc. of NeurIPS (2016)
-
-\bibitem{Hren2008}
-Hren, J.F., Munos, R.: {Optimistic planning of deterministic systems}. Lecture
-  Notes in Computer Science  (2008)
-
-\bibitem{Kearns2002}
-Kearns, M., Mansour, Y., Ng, A.Y.: {A sparse sampling algorithm for
-  near-optimal planning in large Markov decision processes}. In: Proc. of IJCAI
-  (2002)
-
-\bibitem{Kocsis2006}
-Kocsis, L., Szepesv\'{a}ri, C.: Bandit based monte-carlo planning. In: Proc. of
-  ECML (2006)
-
-\bibitem{Leurent2018}
-Leurent, E., Blanco, Y., Efimov, D., Maillard, O.A.: {Approximate Robust
-  Control of Uncertain Dynamical Systems}. NeurIPS Machine Learning for
-  Intelligent Transportation Systems Workshop  (2018)
-
-\bibitem{Munos2014}
-Munos, R.: {From Bandits to Monte-Carlo Tree Search: The Optimistic Principle
-  Applied to Optimization and Planning}. Foundations and
-  Trends{\textregistered} in Machine Learning  (2014)
-
-\bibitem{Szorenyi2014}
-Szorenyi, B., Kedenburg, G., Munos, R.: {Optimistic planning in Markov decision
-  processes using a generative model}. In: Proc. of NeurIPS (2014)
-
-\end{thebibliography}
+\bibliography{references}
 

### 2019-06-25 paper/poster/poster.tex
Updates from Overleaf

@@
 We consider the optimal control of an MDP $\cM = (\cS, cA, R, T, \gamma)$ with bounded rewards $R$
-\begin{itemize}
-    \item $R$ and $T$ are \hlr{unknown}
-    \item Access to a \hlb{generative model} $s' \sim \probability{s'|s,a}$ and $r\sim\probability{r|s, a}$
-    \item \hlr{Fixed-budget} setting: the generative model can be queried $n$ times
+\begin{itemize}[nolistsep]
+    \item[\incarrow] $R$ and $T$ are \hlr{unknown}
+    \item[\incarrow] Access to a \hlb{generative model} $s' \sim \probability{s'|s,a}$ and $r\sim\probability{r|s, a}$
+    \item[\incarrow] \hlr{Fixed-budget} setting: the generative model is \hlr{costly}, can only be queried $n$ times
 \end{itemize}
@@
 
-Split the budget $n$ into $M$ sequences of length $L$.
+\begin{center}
+    \hlb{Optimism in the Face of Uncertainty}\\
+    \incarrow~ in \hlr{observed} and \hlr{future} rewards
+
+\end{center}
+%Split the budget $n$ into $M$ sequences of length $L$.
 %Let $M$ be the largest integer such that $M \log M/(2 \log 1/\gamma) \leq n$\;
@@
 
-Overly pessimistic, especially in the low-budget regime.
+\hlr{Overly pessimistic}, especially in the low-budget regime.
 

### 2019-06-25 paper/main.tex
Include expanded trees in main manuscript

@@
 \begin{itemize}
-    \item \texttt{Random}: returns an action at random, we use it as a minimal performance baseline.
+    \item \texttt{Random}: returns a random action, we use it as a minimal performance baseline.
     \item \texttt{OPD}: the \emph{Optimistic Planning for Deterministic systems} from \citep{Hren2008}, used as a baseline of optimal performance.
@@
 \subfloat{
-	\includegraphics[width=\textwidth]{img/hw_return_svg-tex}
+	\includegraphics[width=0.33\textwidth]{img/hw_return_svg-tex}
 	\label{sub:highway}
 }
-\newline
 \subfloat{
-	\includegraphics[width=0.5\textwidth]{img/gw_return_svg-tex}
+	\includegraphics[width=0.33\textwidth]{img/gw_return_svg-tex}
 	\label{sub:gridworld}
@@
 \subfloat{
-	\includegraphics[width=0.5\textwidth]{img/gw_stoch_return_svf-tex}
+	\includegraphics[width=0.33\textwidth]{img/gw_stoch_return_svf-tex}
 	\label{sub:gridworld_stoch}
@@
 
+\begin{figure}[pht]
+	\centering
+
+	\includegraphics[width=0.43\textwidth]{img/tree-ODP_svg-tex} \includegraphics[width=0.43\textwidth]{img/tree-OLOP_svg-tex}
+	\includegraphics[width=0.43\textwidth]{img/tree-KL-OLOP_svg-tex}
+
+	\caption{The look-ahead trees (down to depth 6) expanded by the planning algorithms from the same initial state in the highway environment with the same budget $n=10^3$.
+The width of edges represents the nodes visit count $T_a(M)$.}
+	\label{fig:trees}
+\end{figure}
+
 The results of our experiments are shown in Figure \ref{fig:experiments}.
@@
 In stark contrast, \KLOLOP makes a much better use of its samples and reaches the same performance an order of magnitude faster.
-Examples of expanded trees are showcased in the Supplementary Material.
+This is illustrated by the expanded trees shown in Figure \ref{fig:trees}: \ODP exploits the deterministic setting and produces a sparse tree densely concentrated around the optimal trajectory.
+Conversely, the tree developed by \OLOP is evenly balanced, which suggests that \OLOP behaves as uniform planning as hypothesised in Section \ref{sec:kl-olop-behaviour}.
+\KLOLOP is more efficient and expands a highly unbalanced tree, exploring the same regions as \ODP.
 Furthermore, in the stochastic gridworld environment shown in Figure \ref{sub:gridworld_stoch}, we observe that the deterministic \texttt{ODP} planner's performance saturates as it settles to suboptimal trajectories, as expected.
@@
 
-\section{Examples of expanded trees}
-
-The trees expanded by different planning algorithms in the Highway environment are displayed in Figure \ref{fig:trees}.
+%\section{Examples of expanded trees}
+%
+%The trees expanded by different planning algorithms in the Highway environment are displayed in Figure \ref{fig:trees}.
 They were all generated in the same initial root state, under the same budget of $n=10^3$ calls to the generative model.
@@
 In stark contrast, \KLOLOP is much more efficient and produced a highly unbalanced tree, exploring the same regions of the tree as \ODP.
-
-\begin{figure}[pht]
-    \centering
-
-    \includegraphics[width=0.64\textwidth]{img/tree-ODP_svg-tex}    \includegraphics[width=0.64\textwidth]{img/tree-OLOP_svg-tex}
-    \includegraphics[width=0.64\textwidth]{img/tree-KL-OLOP_svg-tex}
-
-    \caption{The look-ahead trees expanded by different planning algorithms for a budget $n=10^3$, shown down to depth 6. The width of edges is proportional to the nodes visit count $T_a(M)$.}
-    \label{fig:trees}
-\end{figure}
+%
+%\begin{figure}[pht]
+%    \centering
+%
+%    \includegraphics[width=0.64\textwidth]{img/tree-ODP_svg-tex}    \includegraphics[width=0.64\textwidth]{img/tree-OLOP_svg-tex}
+%    \includegraphics[width=0.64\textwidth]{img/tree-KL-OLOP_svg-tex}
+%
+%    \caption{The look-ahead trees expanded by different planning algorithms for a budget $n=10^3$, shown down to depth 6. The width of edges is proportional to the nodes visit count $T_a(M)$.}
+%    \label{fig:trees}
+%\end{figure}
 

### 2019-08-30 paper/main.tex
Fix typo in image name
Updates from Overleaf

@@
 
-\begin{figure}[htp]
+\begin{figure}[pth]
 \centering
-%\begin{minipage}{\textwidth}
-%      \includegraphics[width=\textwidth]{img/hw_return_svg-tex}
-%      \label{sub:highway}
-%      \includegraphics[width=0.5\textwidth]{img/gw_return_svg-tex}
-%      \label{sub:gridworld}
-%      \includegraphics[width=0.5\textwidth]{img/gw_stoch_return_svg-tex}
-%      \label{sub:gridworld_stoch}
-
 \subfloat{
-	\includegraphics[width=0.33\textwidth]{img/hw_return_svg-tex}
+	\includegraphics[width=\textwidth]{img/hw_return_svg-tex}
 	\label{sub:highway}
 }
+\newline
 \subfloat{
-	\includegraphics[width=0.33\textwidth]{img/gw_return_svg-tex}
+	\includegraphics[width=0.49\textwidth]{img/gw_return_svg-tex}
 	\label{sub:gridworld}
@@
 \subfloat{
-	\includegraphics[width=0.33\textwidth]{img/gw_stoch_return_svf-tex}
+	\includegraphics[width=0.49\textwidth]{img/gw_stoch_return_svg-tex}
 	\label{sub:gridworld_stoch}
@@
 
-\begin{figure}[pht]
+\begin{figure}[pth]
 	\centering
 
-	\includegraphics[width=0.43\textwidth]{img/tree-ODP_svg-tex} \includegraphics[width=0.43\textwidth]{img/tree-OLOP_svg-tex}
-	\includegraphics[width=0.43\textwidth]{img/tree-KL-OLOP_svg-tex}
+	\includegraphics[width=0.64\textwidth]{img/tree-ODP_svg-tex} \includegraphics[width=0.64\textwidth]{img/tree-OLOP_svg-tex}
+	\includegraphics[width=0.64\textwidth]{img/tree-KL-OLOP_svg-tex}
 

### 2019-08-30 paper/poster/poster.tex
Updates from Overleaf

@@
 
-\documentclass[paperwidth=36in, paperheight=48in,portrait,fontscale=0.35, margin=2cm]{baposter}
+\documentclass[a0paper,portrait,fontscale=0.35, margin=2cm]{baposter}
 
@@
 Edouard Leurent, Odalric-Ambrym Maillard
-\vspace{-3\baselineskip}
+\vspace{-4\baselineskip}
 }
@@
 
-We consider the optimal control of an MDP $\cM = (\cS, cA, R, T, \gamma)$ with bounded rewards $R$
+We consider the optimal control of an MDP $\cM = (\cS, A, R, T, \gamma)$ with bounded rewards $R\in[0,1]$
 \begin{itemize}[nolistsep]
@@
 
-\begin{center}
-    \hlb{Optimism in the Face of Uncertainty}\\
-    \incarrow~ in \hlr{observed} and \hlr{future} rewards
-
-\end{center}
 %Split the budget $n$ into $M$ sequences of length $L$.
@@
 %Let $L = \log M / (2 \log 1/\gamma)$\;
-
-\includesvg[width=\textwidth]{img/olop-explain.svg}
-\includesvg[width=\textwidth]{img/olop-explain-2.svg}
+\begin{enumerate}
+    \item Sample $M$ sequences of actions of fixed length $L$
+    \begin{center}
+\includesvg[width=0.8\textwidth]{img/olop-explain.svg}
+\end{center}
+    \item Use the \hlb{return structure} to \hlb{generalise} to unseen sequences
+    \begin{center}
+\includesvg[width=0.8\textwidth]{img/olop-explain-2.svg}
+\end{center}
+\item Be \hlb{Optimistic in the Face of Uncertainty}\\
+    \incarrow~ in \hlr{observed} and \hlr{future} rewards
+\end{enumerate}
 
@@
 
-\includesvg[width=\textwidth]{img/olop-explain-3.svg}
+\begin{center}
+\includesvg[width=0.8\textwidth]{img/olop-explain-3.svg}    
+\end{center}
 
@@
     \begin{equation*}
-        U^{\mu}_a(m) \eqdef \max \left\{q\in I: T_a(m) d(\hat{\mu}_a(m), q) \leq f(m) \right\}
+        U^{\mu}_a(m) \eqdef \max \left\{q\in I: d(\hat{\mu}_a(m), q) \leq \frac{f(m)}{T_a(m)} \right\}
     \end{equation*}
@@
 \headerbox{Experiments}{name=experiments,span=2,column=1,below=time, above=bottom}{ 
-
-\textbf{Expanded trees} for a budget $n=10^3$
+     \includesvg[width=0.33\textwidth]{../img/hw_return}
+     \includesvg[width=0.33\textwidth]{../img/gw_return}
+     \includesvg[width=0.33\textwidth]{../img/gw_stoch_return}
+     \begin{center}
+         \textbf{Average return} over 100 runs –- along with its 95\% confidence interval –- with respect to the available budget $n$
+     \end{center}
 
@@
     \includesvg[width=0.33\textwidth]{../img/tree-KL-OLOP}
-
-\textbf{Average return} over 100 runs –- along with its 95\% confidence interval –- with respect to the available budget $n$
-
-     \includesvg[width=0.33\textwidth]{../img/hw_return}
-     \includesvg[width=0.33\textwidth]{../img/gw_return}
-     \includesvg[width=0.33\textwidth]{../img/gw_stoch_return}
+    \begin{center}
+        \textbf{Expanded trees} for a budget $n=10^3$
+    \end{center}
 }

### 2019-08-30 paper/plot_kl.py
Add talk

@@ -2,7 +2,7 @@ import numpy as np
 import matplotlib.pyplot as plt
 # import seaborn as sns
 # sns.set()
-from rl_agents.utils import bernoulli_kullback_leibler
+from rl_agents.math_utils import bernoulli_kullback_leibler
 
 vkl = lambda p, q: p * np.log(p/q) + (1 - p)*np.log((1-p)/(1-q))
 kl = bernoulli_kullback_leibler

### 2019-09-14 paper/main.tex

@@
 The \texttt{Sparse Sampling} algorithm of \citet{Kearns2002} achieves better worst-case performance, but it is still non-polynomial and doesn't adapt to the structure of the MDP.
-In stark contrast, the \emph{Optimistic Planning for Deterministic systems} (\texttt{OPD}) algorithm considered by \citet{Hren2008} in the case of deterministic transitions and rewards exploits the structure of the cumulative discounted reward to achieve a problem-dependent polynomial bound on sample complexity.
+In stark contrast, the \emph{Optimistic Planning for Deterministic systems} (\OPD) algorithm considered by \citet{Hren2008} in the case of deterministic transitions and rewards exploits the structure of the cumulative discounted reward to achieve a problem-dependent polynomial bound on sample complexity.
 A similar line of work in a deterministic setting is that of \texttt{SOOP} and \texttt{OPC} by \cite{Busoniu2013,Busoniu2018} though they focus on continuous action spaces.
-\texttt{OPD} was later extended to stochastic systems with the \emph{Open-Loop Optimistic Planning} (\OLOP) algorithm introduced by \citet{Bubeck2010} in the open-loop setting: we only consider sequences of actions independently of the states that they lead to.
+\OPD was later extended to stochastic systems with the \emph{Open-Loop Optimistic Planning} (\OLOP) algorithm introduced by \citet{Bubeck2010} in the open-loop setting: we only consider sequences of actions independently of the states that they lead to.
 This restriction in the space of policies causes a loss of optimality, but greatly simplifies the planning problem in the cases where the state space is large or infinite.
@@
     \item \texttt{Random}: returns a random action, we use it as a minimal performance baseline.
-    \item \texttt{OPD}: the \emph{Optimistic Planning for Deterministic systems} from \citep{Hren2008}, used as a baseline of optimal performance.
+    \item \OPD: the \emph{Optimistic Planning for Deterministic systems} from \citep{Hren2008}, used as a baseline of optimal performance.
 This planner is only suited for deterministic environments, and exploits this property to obtain faster rates.
@@
 
-	\includegraphics[width=0.64\textwidth]{img/tree-ODP_svg-tex} \includegraphics[width=0.64\textwidth]{img/tree-OLOP_svg-tex}
-	\includegraphics[width=0.64\textwidth]{img/tree-KL-OLOP_svg-tex}
+	\includegraphics[width=0.64\textwidth]{img/tree_OPD_svg-tex} \includegraphics[width=0.64\textwidth]{img/tree_OLOP_svg-tex}
+	\includegraphics[width=0.64\textwidth]{img/tree_KL-OLOP_svg-tex}
 
@@
 The results of our experiments are shown in Figure \ref{fig:experiments}.
-The \texttt{ODP} algorithm converges very quickly to the optimal return in the two first environments, shown in Figure \ref{sub:highway} and Figure \ref{sub:gridworld}, because it exploits their deterministic nature: it needs neither to estimate the rewards through upper-confidence bounds nor to sample whole sequences all the way from the root when expanding a leaf, which provides a significant speedup.
+The \OPD algorithm converges very quickly to the optimal return in the two first environments, shown in Figure \ref{sub:highway} and Figure \ref{sub:gridworld}, because it exploits their deterministic nature: it needs neither to estimate the rewards through upper-confidence bounds nor to sample whole sequences all the way from the root when expanding a leaf, which provides a significant speedup.
 It can be seen as an oracle allowing to measure the conservativeness of stochastic planning algorithms.
@@
 In stark contrast, \KLOLOP makes a much better use of its samples and reaches the same performance an order of magnitude faster.
-This is illustrated by the expanded trees shown in Figure \ref{fig:trees}: \ODP exploits the deterministic setting and produces a sparse tree densely concentrated around the optimal trajectory.
+This is illustrated by the expanded trees shown in Figure \ref{fig:trees}: \OPD exploits the deterministic setting and produces a sparse tree densely concentrated around the optimal trajectory.
 Conversely, the tree developed by \OLOP is evenly balanced, which suggests that \OLOP behaves as uniform planning as hypothesised in Section \ref{sec:kl-olop-behaviour}.
-\KLOLOP is more efficient and expands a highly unbalanced tree, exploring the same regions as \ODP.
-Furthermore, in the stochastic gridworld environment shown in Figure \ref{sub:gridworld_stoch}, we observe that the deterministic \texttt{ODP} planner's performance saturates as it settles to suboptimal trajectories, as expected.
+\KLOLOP is more efficient and expands a highly unbalanced tree, exploring the same regions as \OPD.
+Furthermore, in the stochastic gridworld environment shown in Figure \ref{sub:gridworld_stoch}, we observe that the deterministic \OPD planner's performance saturates as it settles to suboptimal trajectories, as expected.
 Conversely, the stochastic planners all find better-performing open-loop policies, which justifies the need for this framework.
@@
 \bibliographystyle{splncs04}
-\bibliography{references}
+\bibliography{biblio}
 
@@
 They were all generated in the same initial root state, under the same budget of $n=10^3$ calls to the generative model.
-We observe that \ODP exploits the deterministic setting efficiently and produces a very sparse tree densely concentrated around the optimal trajectory.
+We observe that \OPD exploits the deterministic setting efficiently and produces a very sparse tree densely concentrated around the optimal trajectory.
 Conversely, the tree developed by the \OLOP algorithm is quite evenly balanced, which suggests that \OLOP behaves as uniform planning as hypothesised in \ref{sec:kl-olop-behaviour}.
-In stark contrast, \KLOLOP is much more efficient and produced a highly unbalanced tree, exploring the same regions of the tree as \ODP.
+In stark contrast, \KLOLOP is much more efficient and produced a highly unbalanced tree, exploring the same regions of the tree as \OPD.
 %
@@
 %
-%    \includegraphics[width=0.64\textwidth]{img/tree-ODP_svg-tex}    \includegraphics[width=0.64\textwidth]{img/tree-OLOP_svg-tex}
+%    \includegraphics[width=0.64\textwidth]{img/tree-OPD_svg-tex}    \includegraphics[width=0.64\textwidth]{img/tree-OLOP_svg-tex}
 %    \includegraphics[width=0.64\textwidth]{img/tree-KL-OLOP_svg-tex}

### 2019-09-14 paper/poster/poster.tex

@@
 
-\bibliography{../references.bib}
+\bibliography{../biblio.bib}
 
@@
 \end{itemize}
+\texttt{UCT}: \hlr{doubly-exponential} $\xrightarrow{}$ \OPD: \hlg{polynomial}, \hlr{deterministic}
 }
@@
     \begin{center}
-\includesvg[width=0.8\textwidth]{img/olop-explain.svg}
+\includesvg[width=0.75\textwidth]{img/olop-explain.svg}
 \end{center}
@@
     \begin{center}
-\includesvg[width=0.8\textwidth]{img/olop-explain-2.svg}
+\includesvg[width=0.75\textwidth]{img/olop-explain-2.svg}
 \end{center}
@@
 
-\hlr{Overly pessimistic}, especially in the low-budget regime.
+\hlr{Overly optimistic}, especially in the low-budget regime.
 
     \begin{equation}
-         U^{\mu}_a(m) = \hat{\mu}_a(m) + \sqrt{\frac{2 \log M}{T_a(m)}}
+         \underbrace{U^{\mu}_a(m)}_{\text{Upper bound}} \eqdef \underbrace{\hat{\mu}_a(m)}_{\text{Empirical mean}} + \underbrace{\sqrt{\frac{2 \log M}{T_a(m)}}}_{\text{Confidence interval}}
     \end{equation}
@@
     \label{eq:Ua}
-        U_a(m) \eqdef \sum_{t = 1}^h \gamma^t U^{\mu}_{a_{1:t}}(m) + \frac{\gamma^{h+1}}{1-\gamma}
+        U_a(m) \eqdef \sum_{t=1}^h \underbrace{\gamma^t U^{\mu}_{a_{1:t}}(m)}_{\text{Past rewards}} + \underbrace{\frac{\gamma^{h+1}}{1-\gamma}}_{\text{Future rewards}}
     \end{equation}
@@
     Intuitive explanation: 
-    \begin{itemize}
+    \begin{itemize}[noitemsep]
         \item Unintended behaviour happens when $\hler{U^{\mu}_a(m) > 1}, \forall a$. 
@@
 \begin{center}
-\includesvg[width=0.8\textwidth]{img/olop-explain-3.svg}    
+\includesvg[width=0.75\textwidth]{img/olop-explain-3.svg}    
 \end{center}
@@
 
-    \includesvg[width=0.33\textwidth]{../img/tree-ODP}    \includesvg[width=0.33\textwidth]{../img/tree-OLOP}
-    \includesvg[width=0.33\textwidth]{../img/tree-KL-OLOP}
+    \includesvg[width=0.33\textwidth]{../img/tree_OPD}    \includesvg[width=0.33\textwidth]{../img/tree_OLOP}
+    \includesvg[width=0.33\textwidth]{../img/tree_KL-OLOP}
     \begin{center}
